---
title: "Proteome Discoverer TMT data processing"
subtitle: "`r paste0('Generated with einprot v', packageVersion('einprot'))`"
author:  "Charlotte Soneson (FMI Computational Biology)"
date: "`r Sys.time()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
        css: /tungstenfs/groups/gbioinfo/sonechar/Branding/Rmd_FMI_logo_above_TOC/Rmd_FMI_logo_above_TOC.css
editor_options: 
  chunk_output_type: console
references:
- id: Ritchie2015limma
  title: limma powers differential expression analyses for RNA-sequencing and microarray studies
  author:
  - family: Ritchie
    given: M E
  - family: Phipson
    given: B
  - family: Wu
    given: D
  - family: Hu
    given: Y
  - family: Law
    given: C W
  - family: Shi
    given: W
  - family: Smyth
    given: G K
  container-title: Nucleic Acids Research
  volume: 43
  issue: 7
  page: e47
  type: article-journal
  URL: https://academic.oup.com/nar/article/43/7/e47/2414268
  issued:
    year: 2015
- id: McCarthy2009treat
  title: Testing significance relative to a fold-change threshold is a TREAT
  author:
  - family: McCarthy
    given: D J
  - family: Smyth
    given: G K
  container-title: Bioinformatics
  volume: 25
  page: 765-771
  type: article-journal
  URL: http://bioinformatics.oxfordjournals.org/content/25/6/765
  issued:
    year: 2009
- id: Wu2012camera
  title: Camera:a competitive gene set test accounting for inter-gene correlation
  author:
  - family: Wu
    given: D
  - family: Smyth
    given: G K
  container-title: Nucleic Acids Research
  volume: 40
  issue: 17
  page: e133
  type: article-journal
  URL: https://academic.oup.com/nar/article/40/17/e133/2411151
  issued:
    year: 2012
- id: Phipson2016robust
  title: Robust hyperparameter estimation protects against hypervariable genes and improves power to detect differential expression
  author:
  - family: Phipson
    given: B
  - family: Lee
    given: S
  - family: Majewski
    given: I J
  - family: Alexander
    given: W S
  - family: Smyth
    given: G K
  container-title: Annals of Applied Statistics 
  volume: 10
  page: 946-963
  type: article-journal
  URL: http://projecteuclid.org/euclid.aoas/1469199900
  issued:
    year: 2016
- id: Tusher2001sam
  title: Significance analysis of microarrays applied to the ionizing radiation response
  author:
  - family: Tusher
    given: V G
  - family: Tibshirani
    given: R
  - family: Chu
    given: G
  container-title: Proceedings of the National Academy of Sciences of the United States of America 
  volume: 98
  page: 5116-5121
  type: article-journal
  URL: https://www.pnas.org/content/98/9/5116
  issued:
    year: 2001
- id: Tyanova2016perseus
  title: The Perseus computational platform for comprehensive analysis of (prote)omics data
  author:
  - family: Tyanova
    given: S
  - family: Temu
    given: T
  - family: Sinitcyn
    given: P
  - family: Carlson
    given: A
  - family: Hein
    given: M Y
  - family: Geiger
    given: T
  - family: Mann
    given: M
  - family: Cox
    given: J
  container-title: Nature Methods 
  volume: 13
  page: 731-740
  type: article-journal
  URL: https://www.nature.com/articles/nmeth.3901
  issued:
    year: 2016
- id: Orsburn2021PD
  title: Proteome discoverer-a community enhanced data processing suite for protein informatics 
  author:
  - family: Orsburn
    given: BC
  container-title: J Proteome Res.
  volume: 20(7)
  page: 3497–3507
  type: article-journal
  URL: https://www.mdpi.com/2227-7382/9/1/15
  issued:
    year: 2021
---

{{ConfigParametersStart}}

{{ConfigParametersEnd}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
## Path to Excel file with experiment information
submission_list <- "/group_fmi/ganaly/PAF_RESULTS/SampleSubmissions.xlsx"
```

```{r, echo=FALSE}
## Load packages
suppressPackageStartupMessages({
    library(dplyr)
    library(kableExtra)
    library(tibble)
    library(QFeatures)
    library(SummarizedExperiment)
    library(SingleCellExperiment)
    library(magrittr)
    library(imputeLCMD)
    library(BiocSingular)
    library(scater)
    library(tidyr)
    library(iSEE)
    library(iSEEu)
    library(limma)
    library(ggrepel)
    library(rrcovNA)
    library(circlize)
    library(ComplexHeatmap)
    library(mclust)
    library(DT)
    library(vsn)
    library(readxl)
    library(XML)
    library(UpSetR)
    library(msigdbr)
    library(genefilter)
    library(GSEABase)
})
options(cores = 2)
```

This report describes a reproducible end-to-end analysis of a proteomics dataset quantified 
with [Proteome Discoverer](https://www.thermofisher.com/ch/en/home/industrial/mass-spectrometry/liquid-chromatography-mass-spectrometry-lc-ms/lc-ms-software/multi-omics-data-analysis/proteome-discoverer-software.html) [@Orsburn2021PD]. 
Most of the code is hidden by default, 
but can be displayed by clicking on the `Code` buttons (or by selecting 
`Code -> Show All Code` in the top right corner of the report).

```{r, echo=FALSE, warning=FALSE}
## Read list of submissions and extract the desired one
submission_list <- readxl::read_excel(submission_list, n_max = Inf, 
                                      guess_max = 1e6)
experiment_info <- subset(submission_list, ExperimentID %in% 
                              as.numeric(experimentId))
if (nrow(experiment_info) != 1) {
    stop("You have specified an invalid experiment ID")
}

## Define URL to ProteoLab
Proteolab_Exp_url <- paste0("https://wwwapp.fmi.ch/fileloader/DoRedirect.aspx?e=", 
                            experimentId)

## Extract information about experiment
ei_user <- paste0(experiment_info$OwnerFullName, " (", 
                  experiment_info$OwnerUserName, ")")
ei_group <- experiment_info$GroupName
ei_operator <- experiment_info$Operator
ei_submission_date <- as.character(as.Date(experiment_info$Created))
ei_project_name <- experiment_info$ProjectName
ei_project_id <- experiment_info$ProjectID
ei_experiment_name <- experiment_info$ExperimentName
ei_project_description <- experiment_info$ProjectDescription
ei_experiment_description <- experiment_info$ExperimentDescription
ei_sample_type <- experiment_info$SampleType
ei_nbr_samples <- experiment_info$NoSamples
ei_species <- experiment_info$Species
if (ei_species == "C.elegans") {
    ei_species <- "Caenorhabditis elegans"
}
ei_type_of_analysis <- experiment_info$MasterTemplateName
```

```{r, echo=FALSE}
## Define taxonomic table of supported species
taxTable <- data.frame(
    taxId = c(10090, 9606, 6239, 7955, 7227, 4932, 284812),
    species = c("Mus musculus", "Homo sapiens", "Caenorhabditis elegans",
                "Danio rerio", "Drosophila melanogaster", "Saccharomyces cerevisiae",
                "Schizosaccharomyces pombe 972h-"),
    species_common = c("mouse", "human", "roundworm", "zebrafish", 
                       "fruitfly", "baker's yeast", "fission yeast")
)

## Extract the species_id, common species name and taxonomy ID for later use
if (tolower(ei_species) %in% tolower(taxTable$species_common)) {
    species_id <- taxTable$species[match(tolower(ei_species),
                                         tolower(taxTable$species_common))]
    species_common <- taxTable$species_common[match(tolower(ei_species),
                                                    tolower(taxTable$species_common))]
} else if (tolower(ei_species) %in% tolower(taxTable$species)) {
    species_id <- taxTable$species[match(tolower(ei_species),
                                         tolower(taxTable$species))]
    species_common <- taxTable$species_common[match(tolower(ei_species),
                                                    tolower(taxTable$species))]
} else {
    stop("Unknown species. Information about the species was extracted ",
         "from the sample submissions file ", 
         "(/group_fmi/ganaly/PAF_RESULTS/SampleSubmissions.xlsx).")
}
tax_id <- taxTable$taxId[match(species_id, taxTable$species)]
```

```{r, echo=FALSE}
## Define path to complex DB
if ("complexes" %in% includeFeatureCollections) {
    ## Currently only supported for human, mouse, yeast and worm
    if (species_common %in% c("human", "mouse", "baker's yeast", 
                              "fission yeast", "roundworm")) {
        complexes_file <- complexDbPath
    } else {
        stop("'complexes' can only be tested for human, mouse, yeast and roundworm")
    }
} else {
    complexes_file <- ""
}
```

```{r, echo=FALSE}
## Read PD WorkFlow XML files
pd_PWF <- xmlToList(xmlParse(pdPWFFile))
pd_InputFiles <- read.delim(
    file.path(pdOutputFolder, 
              paste0(pdResultName,  "_InputFiles.txt")), sep = "\t")
pd_StudyInformation <- read.delim(
    file.path(pdOutputFolder, 
              paste0(pdResultName, "_StudyInformation.txt")), sep = "\t")
pd_version <- sub("Created with Discoverer version: ", "",
                  pd_InputFiles$Software.Revision[1])
pd_instruments <- unique(pd_InputFiles$Instrument.Name[-1])
pd_raw_dirs <- ifelse(
    length(unique(gsub("(.+)\\\\.*.raw", "\\1", 
                       pd_InputFiles$File.Name[grep(".raw", pd_InputFiles$File.Name)],
                       perl = TRUE))) == 1, 
    gsub("\\\\", "/", unique(gsub("(.+\\\\).*.raw", "\\1",
                                  pd_InputFiles$File.Name[grep(".raw",
                                                               pd_InputFiles$File.Name)],
                                  perl = TRUE))),
    "multiple")
pd_raw_files <- paste(gsub(".+\\\\(.+)", "\\1", 
                           pd_InputFiles$File.Name[grep(".raw",
                                                        pd_InputFiles$File.Name)]),
                      collapse = ", ")
pd_search_results <- paste(gsub(".+\\\\(.+)", "\\1", 
                                pd_InputFiles$File.Name[grep(".msf",
                                                             pd_InputFiles$File.Name)]),
                           collapse = ", ")
pd_experiments <- paste(unique(pd_StudyInformation$Sample.Group), collapse = ", ")
## TODO: Where is this specified?
pd_calibration <- FALSE
pd_fasta_files <- "D:/Data/seebajan/FASTA/MOUSE__210503.fasta"
pd_contaminants <- "D:/Data/seebajan/FASTA/CON_iRT_contaminants_cRAPMaxQFMI_150507.fasta" 

# mq_fasta_files <- find in pd_Analysis file, look for 
# <ProcessingNodeParameters>
#   <ProcessingNodeParameter  Name="ContaminantsDatabase"

pd_search_engine <- "Sequest HT"
pd_enzymes <- "Trypsin [full]"
pd_fixed_modifications <- "tbd"
pd_variable_modifications <- "tbd"

#### check what can be extracted from .pd_Analysis file
# pd_Analysis$AnalysisSequence$SequenceStep$PrecedingSteps$SequenceStep$WorkflowInfo$WorkflowDefinition$Workflow$WorkflowTree$WorkflowNode

pd_PSM_validation <- "Percolator"
pd_quant_mode <- "Reporter Ions Quantifier"
pd_peptides_for_quantification <- "Peptides + Razor"
pd_quant_order <- "MS2"
pd_abundance <- "S/N"
pd_quanvaluecorrection <- "TRUE"
pd_CoIsolationThr <- 50
pd_avReporterSNThr <- 0
pd_SPSMMpct <- 65
pd_normMode <- "TotalPeptide"
pd_ImputationMode <- "None"
pd_ProtMarker <- "Contaminants, MOUSE"
pd_ConfidenceThr <- 0.01
pd_SearchSettings <- paste0()
pd_quant_methods <- paste0("Peptides used:", pd_peptides_for_quantification,
                           ", quan. method: ", pd_quant_mode,  
                           ", quan. MS order: ", pd_quant_order, 
                           ", abundance type: ", pd_abundance, 
                           ", quan. correction: ", pd_quanvaluecorrection, 
                           ", MS1 co-isolation threshold: ", pd_CoIsolationThr, 
                           ", av. reporter SN threshold: ", pd_avReporterSNThr, 
                           ", PSP mass matches [%]: ", pd_SPSMMpct,
                           ", norm. method: ", pd_normMode,
                           ", PD imputation: ", pd_ImputationMode)

pdFile <- file.path(pdOutputFolder, paste0(pdResultName, "_Proteins.txt"))
```

# Experiment submission details

```{r, echo=FALSE}
tbl <- data.frame(
    c2 = c("User" = ei_user, 
           "Group" = ei_group, 
           "Operator" = ei_operator, 
           "Submission date" = ei_submission_date,
           "Project name" = ei_project_name, 
           "Project ID" = ei_project_id, 
           "Experiment name" = ei_experiment_name,
           "Experiment ID" = experimentId,
           "Project description" = ei_project_description, 
           "Experiment description" = ei_experiment_description, 
           "Type of analysis" = ei_type_of_analysis,
           "Sample type" = ei_sample_type,
           "Number of samples" = ei_nbr_samples,
           "Analysis details" = analysisDetails, 
           "Species" = species_id,
           "Species (common)" = species_common,
           "Taxonomic ID" = tax_id,
           "Cys alkylation" = cysAlkylation,
           "Sample is" = sampleIs,
           "Enzymes" = enzymes)
) %>%
    tibble::rownames_to_column("c1") %>%
    setNames(NULL)

kableExtra::kbl(tbl) %>%
    kableExtra::kable_paper(full_width = TRUE, lightable_options = c("striped"),
                            html_font = "\"Trebuchet MS\", verdana, sans-serif") %>%
    kableExtra::column_spec(column = 1, width = "25em", bold = TRUE, 
                            border_right = TRUE) %>%
    kableExtra::column_spec(column = 2, width = "100em")
```

## Summary

See Report (Rpt_), experimental design (Exd_) and additional info in 
[ProteoLab](`r Proteolab_Exp_url`){target="_blank"}.

## MaxQuant analysis summary

```{r, echo=FALSE}
pd_tbl <- data.frame(
    c2 = c("PD version" = pd_version, 
           "PD result file" = pdResultName,
           "PD analysis file" = pdAnalysisFile,
           "PD Processing WF" = basename(pdPWFFile), 
           "PD Consensus WF" = basename(pdCWFFile), 
           "Search engine" = pd_search_engine,
           "Raw file location" = pd_raw_dirs,
           "Raw files" = pd_raw_files,
           "Sample names" = pd_experiments,
           "Databases" = pd_fasta_files,
           "Contaminants" = pd_contaminants,
           "Quantification settings (LFQ)" = pd_quant_methods,
           "Enzymes" = pd_enzymes, 
           "Variable modifications" = pd_variable_modifications,
           "Fixed modifications" = pd_fixed_modifications)
) %>%
    tibble::rownames_to_column("c1") %>%
    setNames(NULL)

kableExtra::kbl(pd_tbl) %>%
    kableExtra::kable_paper(full_width = TRUE, lightable_options = c("striped"),
                            html_font = "\"Trebuchet MS\", verdana, sans-serif") %>%
    kableExtra::column_spec(column = 1, width = "25em", bold = TRUE, 
                            border_right = TRUE) %>%
    kableExtra::column_spec(column = 2, width = "100em")
```

# Settings {#settings-table}

The following settings define the analysis that will be performed below.

```{r, echo = FALSE}
tbl <- data.frame(
    c2 = c("Name of base assay" = aName,
           "Column pattern" = iColPattern,
           "Sample name pattern" = samplePattern,
           "Include only samples (if applicable)" = paste(includeOnlySamples, 
                                                          collapse = ", "),
           "Exclude samples" = paste(excludeSamples, collapse = ", "),
           "Min. number of peptides" = minPeptides,
           "Min. protein score" = minScore,
           "Imputation method" = imputeMethod,
           "All pairwise comparisons" = allPairwiseComparisons,
           "Min. nbr valid values" = minNbrValidValues,
           "Comparisons" = paste(unlist(lapply(comparisons, 
                                               function(x) paste(x, collapse = " vs "))),
                                 collapse = "; "),
           "Control group" = ctrlGroup,
           "Do all pairwise comparisons" = allPairwiseComparisons,
           "Normalization method" = normMethod,
           "Statistical test" = stattest,
           "Min nbr of valid values to run test" = minNbrValidValues,
           "Minimal fold change (limma/treat)" = minlFC,
           "Adjusted p-value threshold for volcano plots" = volcanoAdjPvalThr,
           "Log2 FC threshold for volcano plots" = volcanoLog2FCThr,
           "Max nbr features to indicate in volcano plots" = volcanoMaxFeatures,
           "s0" = volcanoS0,
           "Features to always label in volcano plots" = paste(volcanoFeaturesToLabel,
                                                               collapse = ", "),
           "Feature collections" = paste(includeFeatureCollections, collapse = "; "),
           "Complexes file" = gsub(".+\\/(.+.rds)", "\\1", complexes_file),
           "Complexes from species" = complexSpecies,
           "Custom complexes" = names(customComplexes),
           "FDR Threshold for complexes" = complexFDRThr,
           "Number of permutations" = nperm,
           "Random seed" = seed)
) %>%
    tibble::rownames_to_column("c1")

if (stattest == "ttest") {
    tbl <- tbl %>% 
        dplyr::filter(!(c1 %in% c("Minimal fold change (limma/treat)",
                                  "Log2 FC threshold for volcano plots")))
}
if (stattest == "limma") {
    tbl <- tbl %>%
        dplyr::filter(!(c1 %in% c("s0", "Number of permutations")))
}
tbl <- tbl %>% setNames(NULL)

kableExtra::kbl(tbl) %>%
    kableExtra::kable_paper(full_width = TRUE, lightable_options = c("striped"),
                            html_font = "\"Trebuchet MS\", verdana, sans-serif") %>%
    kableExtra::column_spec(column = 1, width = "100em", bold = TRUE, 
                            border_right = TRUE) %>%
    kableExtra::column_spec(column = 2, width = "100em")
```

# Prepare feature collections for later testing

In addition to testing individual proteins for differential abundance 
between groups, we can also test collections of proteins. Here we 
define the collections that will be used. 

```{r}
## Help function to replace every nth ";" by "; "
createPattern <- function(n) {
    sprintf("(%s[^;]+);", strrep("[^;]+;", n - 1))
}
pat <- createPattern(10)

featureCollections <- list()

## Complexes
## -----------------------------------------------------------------
if ("complexes" %in% includeFeatureCollections) {
    complexes <- readRDS(complexes_file)
    if (species_common %in% names(complexes)) {
        crl <- complexes[[species_common]]
    } else if (species_id %in% names(complexes)) {
        crl <- complexes[[species_id]]
    } else {
        stop("No complex database available for the current species")
    }
    if (complexSpecies == "current") {
        ## Only test complexes defined for the current species
        crl <- crl[mcols(crl)$Species.common %in% 
                       c(species_id, species_common)]
    }
    mcols(crl)$genes <- sapply(crl, 
                               function(w) gsub(pat, "\\1; ", 
                                                paste(w, collapse = ";")))
} else {
    crl <- CharacterList()
}

## Custom complexes (to add to the list of complexes above)
if (length(customComplexes) != 0) {
    tmpcompl <- as(customComplexes, "CharacterList")
    mcols(tmpcompl) <- DataFrame(
        Species.common = species_common, 
        Source = "custom",
        PMID = NA_character_,
        All.names = names(tmpcompl),
        genes = sapply(customComplexes, 
                       function(w) gsub(pat, "\\1; ",
                                        paste(w, collapse = ";")))
    )
    if (length(crl) > 0) {
        crl <- c(crl, tmpcompl)
    } else {
        crl <- tmpcompl
    }
}

if (length(crl) > 0) {
    featureCollections$complexes <- crl
}

## GO terms
## -----------------------------------------------------------------
if ("GO" %in% includeFeatureCollections) {
    goannots <- msigdbr::msigdbr(species = species_id, category = "C5") %>%
        dplyr::select(gs_name, gene_symbol)
    goannots <- as(lapply(split(goannots, f = goannots$gs_name), 
                          function(w) unique(w$gene_symbol)),
                   "CharacterList")
    mcols(goannots)$genes <- sapply(goannots, function(w) 
        gsub(pat, "\\1; ", paste(w, collapse = ";"))
    )

    featureCollections$GO <- goannots
}

featureCollections
```

# Read PD output

The input to this workflow is a `Proteins.txt` file from Proteome Discoverer 
(see path in the table above). We read the PD intensities into `R` 
and store them in a 
[QFeatures](https://bioconductor.org/packages/QFeatures/) object. This object 
will later be expanded with additional data, such as transformed and 
imputed quantifications. 

Based on the 'Column pattern' indicated in the table above, the following columns 
are determined to contain the intensity values of interest:

```{r}
## Columns representing intensities
(iColsAll <- grep(iColPattern, names(read.delim(pdFile, nrows = 2)), value = TRUE))
```

```{r}
if (length(iColsAll) == 0) {
    stop("No samples were found matching the specified iColPattern.")
}
```

After any explicit selection of samples to include or exclude, as specified 
above, the following samples are retained and will be available for use in the 
downstream analysis: 

```{r}
if (length(includeOnlySamples) > 1 || includeOnlySamples != "") {
    ## Specify samples to include
    iCols <- iColsAll[!is.na(stringr::str_extract(
        iColsAll, paste(includeOnlySamples, collapse = "|")))]
} else if (length(excludeSamples) > 1 || excludeSamples != "") {
    ## Specify samples to exclude
    iCols <- iColsAll[is.na(stringr::str_extract(
        iColsAll, paste(excludeSamples, collapse = "|")))]
} else {
    iCols <- iColsAll
}
iCols
```

At this point, the QFeatures object holds an _assay_ with the intensities from 
PD, as well as annotations corresponding to all other columns in the 
PD file. The assay name will be `r aName`, as specified via the 
`aName` variable above. 

```{r}
if (length(iCols) == 0) {
    stop("No samples were retained - please check the specification ",
         "of includeOnlySamples/excludeSamples (note that these should ",
         "specify individual samples, not group names).")
}

## Read Proteome Discoverer output
qft <- QFeatures::readQFeatures(pdFile, ecol = iCols, name = aName, 
                                sep = "\t")
qft
```

# Modify feature names

The feature ID used when reading the data above are the numeric indices provided in the 
'id' column. We replace these IDs with more interpretable ones, corresponding to 
the gene symbol, or, in cases where that is missing, with 
the accession. 

```{r}
## Extract the first annotated gene name
gName <- vapply(strsplit(rowData(qft[[aName]])$Gene.Symbol, "; "), 
                .subset, 1, FUN.VALUE = "NA")

## Extract the accession
accession <- vapply(strsplit(rowData(qft[[aName]])$Accession, "; "), 
                    .subset, 1, FUN.VALUE = "NA")

## Generate IDs for STRING
stringIDs <- gName
idxna <- which(is.na(stringIDs))
stringIDs[idxna] <- accession[idxna]
rowData(qft[[aName]])$IDsForSTRING <- stringIDs

## If there are duplicated gene IDs, make them unique by appending the 
## respective accession ID
idxdup <- which(duplicated(gName) | is.na(gName))
idxdup <- which(gName %in% gName[idxdup])
gName[idxdup] <- paste0(gName[idxdup], ".", accession[idxdup])

## Make sure that there are no duplicated IDs and set as row names
gName <- make.unique(gName, sep = ".")
stopifnot(all(!duplicated(gName)))
rownames(qft[[aName]]) <- gName

## Modify feature collections so that they 
## contain any row name corresponding to the features in the collection.
## Some rows in the data set don't correspond to any gene name 
## (and are represented by their majority protein ID).
## Other rows correspond to multiple gene names (and are represented 
## by the first of these). 
## Finally, there may be multiple rows corresponding to the same main gene
## (in which case a majority protein ID is appended to the row name).
## Currently, we consider the last group, and expand feature 
## collections to include all variants of the annotated gene.
allGeneNames <- strsplit(rowData(qft[[aName]])$Gene.Symbol, "; ")
names(allGeneNames) <- rownames(qft[[aName]])
## Only consider rows corresponding to a single gene name
allGeneNames <- allGeneNames[sapply(allGeneNames, length) == 1]
## Since each element has length 1, we can simplify
allGeneNames <- unlist(allGeneNames)
featureCollections <- lapply(featureCollections, function(fcoll) {
    for (nm in names(fcoll)) {
        fcoll[[nm]] <- unique(c(fcoll[[nm]], 
                                names(allGeneNames)[allGeneNames %in% fcoll[[nm]]]))
    }
    fcoll
})
```

# Add sample annotations

Next, we compile the sample annotations. The sample ID and group annotation 
will be extracted from the sample name in the `MaxQuant` file. The `group` 
column will be used to define groups for the statistical testing later. 
Please check that the table below correspond to your expectations.

```{r}
## Get sample ID by removing the iColPattern from the colnames
cdn <- sub(iColPattern, "", colnames(qft)[[aName]])
qft$sample <- cdn

## Get group ID by removing the samplePattern from the sample ID
qft$group_orig <- sub(samplePattern, "", cdn)

## Define a new grouping, where all samples from the groups designed as
## control groups are assigned to the same group (to be used as the 
## baseline for the statistical tests when applicable)
qft$group <- qft$group_orig
if (length(comparisons) == 0) {
    if ((allPairwiseComparisons && !all(setdiff(ctrlGroup, "") %in% qft$group)) ||
        (!allPairwiseComparisons && !all(ctrlGroup %in% qft$group))) {
        stop("Misspecified 'ctrlGroup'")
    }
}
mergedCtrlGroup <- paste(ctrlGroup, collapse = ".")
qft$group[qft$group %in% ctrlGroup] <- mergedCtrlGroup
ctrlGroup <- mergedCtrlGroup

DT::datatable(as.data.frame(colData(qft)),
              options = list(scrollX = TRUE, pageLength = 20))
```

# Overall distribution of intensities

The box plot below displays the distribution of raw intensities in each sample, 
on a log scale (excluding any missing values).

```{r, warning=FALSE, message=FALSE}
ggplot(as.data.frame(assay(qft[[aName]])) %>%
           tidyr::gather(key = "col_id", value = "intensity") %>%
         dplyr::left_join(as.data.frame(colData(qft)) %>%
                            tibble::rownames_to_column("col_id"),
                          by = "col_id"),
       aes(x = sample, y = intensity, fill = group)) + 
    geom_boxplot(alpha = 0.5) + theme_bw() + 
    scale_y_log10() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
    labs(x = "", y = "Raw intensity")
```


# Filter out contaminants, reverse hits, and proteins with low confidence

Next, we filter out any features classified by MaxQuant as potential 
contaminants or reverse (decoy) hits, and features identified only by site (which 
removes proteins that are only identified by peptides carrying one or more 
modified amino acids). In addition, we may remove protein identifications 
based on score and number of peptides (i.e. to exclude one-hit wonders). 

```{r}
nbrFeaturesBefore <- nrow(qft[[aName]])
filtdf <- as.data.frame(rowData(qft[[aName]])) %>%
    dplyr::select(Contaminant) %>%
    dplyr::mutate(across(c(Contaminant), 
                         function(x) as.numeric(x == "True")))# %>%
    # dplyr::mutate(Score = as.numeric(Score < minScore),
    #               Peptides = as.numeric(Peptides < minPeptides)) 
qft <- qft %>%
    filterFeatures(~ Contaminant == "False")
nbrFeaturesAfter <- nrow(qft[[aName]])

## Add a column to feature collections indicating how many genes 
## are shared with the filtered data set
featureCollections <- lapply(featureCollections, function(fc) {
    mcols(fc)$nSharedGenes <- sapply(
      fc, function(x) length(intersect(x, rownames(qft[[aName]])))
    )
    mcols(fc)$sharedGenes <- sapply(
      fc, function(x) paste(intersect(x, rownames(qft[[aName]])),
                            collapse = ";")
    )
    fc
})

# UpSetR::upset(filtdf)
```

This filtering removed `r nbrFeaturesBefore - nbrFeaturesAfter` features. 
The new `qft` object has `r nbrFeaturesAfter` features.


# Apply a log2 transformation

Before the downstream analysis, we log2-transform the measured intensities. We also 
add an additional assay to keep track of the position of the missing values
(which will be imputed later). The `qft` object now has three assays:

```{r}
qft <- logTransform(qft, base = 2,
                    i = aName,
                    name = paste0("log2_", aName))
qft <- logTransform(qft, base = 2,
                    i = aName,
                    name = paste0("log2_", aName, "_withNA"))

## Add assay indicating missing values, which will be imputed
tmp <- qft[[paste0("log2_", aName)]]
assay(tmp) <- !is.finite(assay(tmp))
qft <- addAssay(qft, tmp,
                name = paste0("imputed_", aName))
qft

## Set the assay to use for tests later (depending on the downstream steps,
## this may be updated below)
assayForTests <- paste0("log2_", aName)
```

# Visualize missing value patterns

The plot below shows the fraction of the total set of features that are detected 
(with a non-missing value) in each of the samples.

```{r}
## Replace zeros/-Inf values by explicit NA values in the assays
qft <- zeroIsNA(qft, aName)
qft <- infIsNA(qft, paste0("log2_", aName))
qft <- infIsNA(qft, paste0("log2_", aName, "_withNA"))

## Count number of NA values
nbr_na <- nNA(qft, i = seq_along(qft))

ggplot(as.data.frame(nbr_na$nNAcols) %>% dplyr::filter(assay == aName),
       aes(x = name, y = 100 - pNA, label = paste0(round(100 - pNA, 1), "%"))) + 
    geom_bar(stat = "identity") + theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
    labs(x = "", y = "Detected features (%)") + 
    expand_limits(y = 100) + 
    geom_text(vjust = 1.5, color = "white", size = 3)
```

We also plot the number of features that are detected in a given number 
of samples.

```{r}
ggplot(as.data.frame(nbr_na$nNArows) %>% dplyr::filter(assay == aName) %>%
           dplyr::group_by(nNA) %>% tally(),
       aes(x = ncol(qft[[aName]]) - nNA, y = n)) + 
    geom_bar(stat = "identity") + theme_bw() + 
    labs(x = "Number of samples in which a feature is detected", 
         y = "Number of features")
```

Finally we show the full missing value structure in the data matrix. The white 
values correspond to missing values, and the grey ones to observed intensities.

```{r, fig.height=7, message=FALSE}
# col_fun = colorRamp2(c(0, 1), c("grey50", "white"))
# Heatmap(assay(qft[[paste0("imputed_", aName)]]) + 0, col = col_fun,
#         name = "imputed", column_title = "Missing value pattern (white = missing)",
#         cluster_rows = TRUE, cluster_columns = TRUE, show_row_names = FALSE,
#         show_heatmap_legend = FALSE, use_raster = TRUE)
```


# Imputation

Next, we apply the `r imputeMethod` method to perform imputation 
of the log2-transformed data.

```{r, eval=(imputeMethod=="MinProb"), echo=(imputeMethod=="MinProb"), results="hide"}
set.seed(seed)
qft <- impute(qft, method = "MinProb", i = paste0("log2_", aName))
```

```{r, eval=(imputeMethod=="impSeqRob"), echo=(imputeMethod=="impSeqRob")}
set.seed(seed)
tmp <- impSeqRob(assay(qft[[paste0("log2_", aName)]]))
assay(qft[[paste0("log2_", aName)]]) <- tmp$x
```

The following histograms show the distribution of imputed and non-imputed 
values in each sample. 

```{r, fig.width=7, fig.height=7, message = FALSE}
plotdf <- as.data.frame(assay(qft[[paste0("log2_", aName)]])) %>%
    tibble::rownames_to_column("pid") %>%
    tidyr::gather(key = "sample", value = "log2intensity", -pid) %>%
    dplyr::left_join(
        as.data.frame(assay(qft[[paste0("imputed_", aName)]])) %>%
            tibble::rownames_to_column("pid") %>%
            tidyr::gather(key = "sample", value = "imputed", -pid)
    ) %>%
    dplyr::mutate(sample = sub(iColPattern, "", sample))
ggplot(plotdf, aes(x = log2intensity, fill = imputed)) + 
    geom_histogram(bins = 50) + facet_wrap(~ sample) + 
    theme_bw() + labs(x = "log2 intensity") + 
    scale_fill_manual(values = c(`TRUE` = "grey", `FALSE` = "firebrick1"))
```

# Overall distribution of log2-intensities

Next we consider the overall distribution of log2-intensities among the 
samples (after imputation).

```{r}
ggplot(as.data.frame(assay(qft[[paste0("log2_", aName)]])) %>%
           tidyr::gather(key = "col_id", value = "log2intensity") %>%
         dplyr::left_join(as.data.frame(colData(qft)) %>%
                            tibble::rownames_to_column("col_id"),
                          by = "col_id"),
       aes(x = sample, y = log2intensity, fill = group)) + 
    geom_boxplot(alpha = 0.5) + theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
    labs(x = "", y = "log2 intensity (after imputation)")
```

We also plot the mean-variance relationship of log2-intensities (after imputation).

```{r, message=FALSE}
ggplot(as.data.frame(assay(qft[[paste0("log2_", aName)]])) %>%
           tibble::rownames_to_column("pid") %>%
           tidyr::gather(key = "sample", value = "log2intensity", -pid) %>%
           dplyr::group_by(pid) %>%
           dplyr::mutate(mean_log2intensity = mean(log2intensity),
                         sd_log2intensity = sd(log2intensity)),
       aes(x = mean_log2intensity, y = sd_log2intensity)) + 
    geom_point(alpha = 0.05) + geom_smooth() + 
    theme_bw() + 
    labs(x = "Mean log2 intensity (after imputation)", 
         y = "SD of log2 intensity (after imputation)")
```


```{r, results="asis", echo=FALSE, eval=(normMethod == "none")}
cat(paste0("The log2 intensities are not normalized further across samples ", 
           "since 'normMethod' is set to 'none'."))
```

```{r, results="asis", echo=FALSE, eval=(normMethod != "none")}
cat(paste0("The log2 intensities are next normalized across samples ", 
           "using the ", normMethod, " method."))
```

```{r, echo=(normMethod != "none"), eval=(normMethod != "none"), message=FALSE}
qft <- normalize(qft,
                 i = paste0("log2_", aName),
                 name = paste0("log2norm_", aName),
                 method = normMethod)

## If data is normalized, use normalized values for testing
assayForTests <- paste0("log2norm_", aName)

ggplot(as.data.frame(assay(qft[[paste0("log2norm_", aName)]])) %>%
           tidyr::gather(key = "col_id", value = "log2intensity") %>%
         dplyr::left_join(as.data.frame(colData(qft)) %>%
                            tibble::rownames_to_column("col_id"),
                          by = "col_id"),
       aes(x = sample, y = log2intensity, fill = group)) + 
    geom_boxplot(alpha = 0.5) + theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
    labs(x = "", y = "log2 intensity (after imputation and normalization)")

ggplot(as.data.frame(assay(qft[[paste0("log2norm_", aName)]])) %>%
           tibble::rownames_to_column("pid") %>%
           tidyr::gather(key = "col_id", value = "log2intensity", -pid) %>%
           dplyr::group_by(pid) %>%
           dplyr::mutate(mean_log2intensity = mean(log2intensity),
                         sd_log2intensity = sd(log2intensity)),
       aes(x = mean_log2intensity, y = sd_log2intensity)) + 
    geom_point(alpha = 0.05) + geom_smooth() + 
    theme_bw() + 
    labs(x = "Mean log2 intensity (after imputation and normalization)", 
         y = "SD of log2 intensity (after imputation and normalization)")
```

# Statistical testing

For each feature, we then compare the (possibly imputed) log2 intensities 
between groups. 

```{r, results="asis", echo=FALSE, eval=(stattest == "limma")}
cat(paste0("For this, we use the treat function from the ", 
           "[limma](https://bioconductor.org/packages/limma/) R/Bioconductor ",
           "package [@McCarthy2009treat; @Ritchie2015limma; @Phipson2016robust]. ",
           "For more information about the df.prior, representing the amount of ", 
           "extra information that is borrowed from the full set of features ", 
           "in order to improve the inference for each feature, see section 13.2 in the ",
           "[limma user guide](https://www.bioconductor.org/packages/devel/bioc/vignettes/limma/inst/doc/usersguide.pdf). ",
           "In addition to the feature-wise tests, we apply the ",
           "camera method [@Wu2012camera] to test for significance of each ", 
           "included feature collection. These tests are based on the ",
           "t-statistic returned from limma."))
```

```{r, results="asis", echo=FALSE, eval=(stattest == "ttest")}
cat(paste0("For this, we use a Student's t-test. To determine which features ",
           "show significant changes, we calculate the SAM statistic [@Tusher2001sam], ",
           "and estimate the false discovery rate at different thresholds ",
           "using permutations, mimicking the approach used by Perseus [@Tyanova2016perseus].",
           "In addition to the feature-wise tests, we apply the ",
           "camera method [@Wu2012camera] to test for significance of each ", 
           "included feature collection. These tests are based on the ",
           "SAM statistic calculated from the t-statistic and the specified S0."))
```

The `r assayForTests` assay will be used for the tests. The 
following pairwise comparisons will be performed (in each case, the 
first listed group will be the 'baseline' group): 

```{r}
## Initialize list for test results and plot titles/notes
tests <- list()
plottitles <- list()
plotnotes <- list()
curveparams <- list()

## Get list of comparisons to make
if (length(comparisons) == 0) {
    ## No manually defined comparisons
    if (allPairwiseComparisons) {
        comparisons <- combn(unique(qft$group), 2, simplify = FALSE)
        comparisons <- lapply(comparisons, function(cmp) {
            if (ctrlGroup %in% cmp && ctrlGroup != cmp[1]) {
                rev(cmp)
            } else {
                cmp
            }
        })
    } else {
        ## Get groups to compare to the control group
        other_groups <- setdiff(unique(qft$group), ctrlGroup)
        comparisons <- lapply(as.list(other_groups), function(x) c(ctrlGroup, x))
    }
}
print(comparisons)
```

```{r, echo=(stattest=="limma"), eval=(stattest=="limma")}
for (cmp in comparisons) {
    ## Subset QFeatures object to only the two groups considered
    qftsub <- qft[, qft$group %in% cmp]
    
    ## limma (treat)
    fc <- factor(qftsub$group, levels = cmp)
    design <- model.matrix(~ fc)
    exprvals <- assay(qftsub[[assayForTests]], withDimnames = TRUE)

    ## Only consider features with at least a given number of valid values
    imputedvals <- assay(qftsub[[paste0("imputed_", aName)]], 
                         withDimnames = TRUE)
    keep <- rowSums(!imputedvals) >= minNbrValidValues
    exprvals <- exprvals[keep, , drop = FALSE]

    fit <- lmFit(exprvals, design)
    fit <- treat(fit, fc = 2^minlFC, trend = TRUE, robust = FALSE)
    res <- topTreat(fit, coef = paste0("fc", cmp[2]), number = Inf, sort.by = "none") %>%
        tibble::rownames_to_column("pid") %>%
        dplyr::mutate(mlog10p = -log10(P.Value))
    
    ## Test feature sets
    featureCollections <- lapply(featureCollections, function(fcoll) {
        camres <- cameraPR(
            statistic = structure(res$t, names = res$pid),
            index = ids2indices(as.list(fcoll), res$pid, remove.empty = FALSE),
            sort = FALSE
        )
        if (!("FDR" %in% colnames(camres))) {
            camres$FDR <- p.adjust(camres$PValue, method = "BH")
        }
        colnames(camres) <- paste0(cmp[2], "_vs_", cmp[1], "_", colnames(camres))
        stopifnot(all(rownames(camres) == names(fcoll)))
        mcols(fcoll) <- cbind(mcols(fcoll), camres)
        fcoll
    })
    
    if ("complexes" %in% names(featureCollections)) {
        tmpres <- as.data.frame(mcols(featureCollections$complexes), optional = TRUE) %>%
            tibble::rownames_to_column("complex") %>%
            dplyr::select(complex, genes, sharedGenes,
                          Source, All.names, PMID,
                          contains(paste0(cmp[2], "_vs_", cmp[1]))) %>%
            dplyr::arrange(.data[[paste0(cmp[2], "_vs_", cmp[1], "_FDR")]]) %>%
            dplyr::filter(.data[[paste0(cmp[2], "_vs_", cmp[1], "_FDR")]] < complexFDRThr)
    #     if (nrow(tmpres) > 0) {
    #         write.table(tmpres, 
    #             file = sub("\\.Rmd$", paste0("_testres_", cmp[2], "_vs_", cmp[1],
    #                                          "_camera_complexes.txt"), 
    #                        knitr::current_input()),
    #             row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
    #     }
    }
    
    plottitles[[paste0(cmp[2], "_vs_", cmp[1])]] <- 
        paste0(cmp[2], " vs ", cmp[1], ", limma treat (H0: |log2FC| <= ", minlFC, ")")
    plotnotes[[paste0(cmp[2], "_vs_", cmp[1])]] <- 
        paste0("df.prior = ", round(fit$df.prior, digits = 2))
    
    res <- data.frame(pid = rownames(imputedvals)) %>%
        dplyr::left_join(res, by = "pid")
    res$showInVolcano <- res$adj.P.Val <= volcanoAdjPvalThr & 
        abs(res$logFC) >= volcanoLog2FCThr
    stopifnot(all(rownames(rowData(qftsub[[aName]])) == res$pid))
    if (iColPattern == "^Abundance\\.") {
        ## If we're using Abundance values for the analysis, they will not be
        ## present in the rowData anymore.
        abundance_values <- as.data.frame(assay(qftsub[[aName]]))
        abundance_values[is.na(abundance_values)] <- 0
    } else {
        abundance_values <- as.data.frame(
            rowData(qftsub[[aName]])[, sub(iColPattern, "Abundance.",
                                           colnames(qftsub[[aName]])), drop = FALSE]
        )
    }
    log_abundance_values <- log2(abundance_values) %>%
        dplyr::mutate(across(everything(), .fns = ~ ifelse(is.finite(.), ., NA)))
    res <- cbind(res, abundance_values)
    for (gr in unique(colData(qftsub)$group)) {
        res[[paste0("Abundance.", gr, ".avg")]] <-
            rowMeans(abundance_values[, colData(qftsub)$group == gr, drop = FALSE])
        res[[paste0("Abundance.", gr, ".sd")]] <-
            genefilter::rowSds(abundance_values[, colData(qftsub)$group == gr, drop = FALSE])
        res[[paste0("log2_Abundance.", gr, ".avg")]] <-
            rowMeans(log_abundance_values[, colData(qftsub)$group == gr, drop = FALSE],
                     na.rm = TRUE)
        res[[paste0("log2_Abundance.", gr, ".sd")]] <-
            genefilter::rowSds(log_abundance_values[, colData(qftsub)$group == gr,
                                               drop = FALSE],
                               na.rm = TRUE)
    }
    tests[[paste0(cmp[2], "_vs_", cmp[1])]] <- res
}
```

```{r, echo=(stattest=="ttest"), eval=(stattest=="ttest")}
for (cmp in comparisons) {
    ## Subset QFeatures object to only the two groups considered
    qftsub <- qft[, qft$group %in% cmp]
    
    ## Student's t-test
    fc <- factor(qftsub$group, levels = rev(cmp))
    exprvals <- assay(qftsub[[assayForTests]], withDimnames = TRUE)
    
    ## Only consider features with at least a given number of valid values
    imputedvals <- assay(qftsub[[paste0("imputed_", aName)]], 
                         withDimnames = TRUE)
    keep <- rowSums(!imputedvals) >= minNbrValidValues
    exprvals <- exprvals[keep, , drop = FALSE]
    
    res <- genefilter::rowttests(exprvals, fac = fc)
    res <- res %>%
        tibble::rownames_to_column("pid") %>%
        dplyr::rename(t = statistic, logFC = dm, P.Value = p.value) %>%
        dplyr::mutate(mlog10p = -log10(P.Value),
                      adj.P.Val = p.adjust(P.Value, method = "BH"),
                      AveExpr = rowMeans(exprvals)) %>% 
        dplyr::mutate(sam = t/(1 + t * volcanoS0/logFC))
    
    ## Test feature sets
    featureCollections <- lapply(featureCollections, function(fcoll) {
        camres <- cameraPR(
            statistic = structure(res$sam, names = res$pid),
            index = ids2indices(as.list(fcoll), res$pid, remove.empty = FALSE),
            sort = FALSE
        )
        if (!("FDR" %in% colnames(camres))) {
            camres$FDR <- p.adjust(camres$PValue, method = "BH")
        }
        colnames(camres) <- paste0(cmp[2], "_vs_", cmp[1], "_", colnames(camres))
        stopifnot(all(rownames(camres) == names(fcoll)))
        mcols(fcoll) <- cbind(mcols(fcoll), camres)
        fcoll
    })
    
    if ("complexes" %in% names(featureCollections)) {
        tmpres <- as.data.frame(mcols(featureCollections$complexes), optional = TRUE) %>%
            tibble::rownames_to_column("complex") %>%
            dplyr::select(complex, genes, sharedGenes,
                          contains(paste0(cmp[2], "_vs_", cmp[1]))) %>%
            dplyr::arrange(.data[[paste0(cmp[2], "_vs_", cmp[1], "_FDR")]]) %>%
            dplyr::filter(.data[[paste0(cmp[2], "_vs_", cmp[1], "_FDR")]] < complexFDRThr)
    #     if (nrow(tmpres) > 0) {
    #         write.table(tmpres, 
    #             file = sub("\\.Rmd$", paste0("_testres_", cmp[2], "_vs_", cmp[1],
    #                                          "_camera_complexes.txt"), 
    #                        knitr::current_input()),
    #             row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
    #     }
    }

    plottitles[[paste0(cmp[2], "_vs_", cmp[1])]] <- 
        paste0(cmp[2], " vs ", cmp[1], ", t-test")
    plotnotes[[paste0(cmp[2], "_vs_", cmp[1])]] <- ""
    
    ## Get the threshold curve (replicating Perseus plots) --------------------
    ## Permute and calculate SAM statistics
    set.seed(seed)
    sampermL <- lapply(seq_len(nperm), function(i) {
        fc0 <- factor(sample(as.character(fc), size = length(fc)))
        ## Don't include the true grouping
        if (mclust::adjustedRandIndex(fc, fc0) != 1) {
            tts0 <- genefilter::rowttests(as.matrix(exprvals), fc0)
            tts0$statistic/(1 + tts0$statistic * volcanoS0/tts0$dm)
        }
    })
    sampermdf <- abs(do.call(cbind, sampermL))
    
    ## Define candidate cutoff values - use the observed SAM statistics plus the 
    ## middle point between each consecutive pair
    cs <- sort(unique(abs(res$sam)))
    cs <- sort(unique(c(cs, cs + c(diff(cs)/2, 0))))
    
    ## Get mean number of FPs across permutations for each cutoff
    nbrFP <- sapply(cs, function(cval) {
        apply(sampermdf, 2, function(w) sum(w >= cval))
    })
    meannfp <- colMeans(nbrFP)
    
    ## Get number of significant hits for each cutoff
    signhits <- sapply(cs, function(cval) {
        sum(abs(res$sam) >= cval)
    })
    
    ## Get all cutoffs that pass the FDR threshold and extract the smallest one
    passthr <- which(meannfp/signhits < volcanoAdjPvalThr)
    if (length(passthr) > 0) {
        ta <- min(cs[passthr])
        ta2 <- cs[min(passthr)]
        lowerbound <- ceiling(100 * ta * volcanoS0)/100
        x <- seq(from = lowerbound, 
                 to = max(lowerbound, 1.1 * max(abs(res$logFC))), by = 0.01)
    } else {
        ta <- Inf
        x <- NA
    }
    df <- sum(table(fc)) - 2
    curveparams[[paste0(cmp[2], "_vs_", cmp[1])]] <- 
        list(x = x, ta = ta, s0 = volcanoS0, df = df)
    
    res <- data.frame(pid = rownames(imputedvals)) %>%
        dplyr::left_join(res, by = "pid")
    res$showInVolcano <- abs(res$sam) >= ta
    stopifnot(all(rownames(rowData(qftsub[[aName]])) == res$pid))
    # if (iColPattern == "^iBAQ\\.") {
    #     ## If we're using iBAQ values for the analysis, they will not be 
    #     ## present in the rowData anymore.
    #     ibaq_values <- as.data.frame(assay(qftsub[[aName]]))
    #     ibaq_values[is.na(ibaq_values)] <- 0
    # } else {
    #     ibaq_values <- as.data.frame(
    #         rowData(qftsub[[aName]])[, sub(iColPattern, "iBAQ.", 
    #                                        colnames(qftsub[[aName]])), drop = FALSE]
    #     )
    # }
    # log_ibaq_values <- log2(ibaq_values) %>%
    #     dplyr::mutate(across(everything(), .fns = ~ ifelse(is.finite(.), ., NA)))
    # res <- cbind(res, ibaq_values)
    # for (gr in unique(colData(qftsub)$group)) {
    #     res[[paste0("iBAQ.", gr, ".avg")]] <- 
    #         rowMeans(ibaq_values[, colData(qftsub)$group == gr, drop = FALSE])
    #     res[[paste0("iBAQ.", gr, ".sd")]] <- 
    #         genefilter::rowSds(ibaq_values[, colData(qftsub)$group == gr, drop = FALSE])
    #     res[[paste0("log2_iBAQ.", gr, ".avg")]] <- 
    #         rowMeans(log_ibaq_values[, colData(qftsub)$group == gr, drop = FALSE], 
    #                  na.rm = TRUE)
    #     res[[paste0("log2_iBAQ.", gr, ".sd")]] <- 
    #         genefilter::rowSds(log_ibaq_values[, colData(qftsub)$group == gr, 
    #                                            drop = FALSE],
    #                            na.rm = TRUE)
    # }
    tests[[paste0(cmp[2], "_vs_", cmp[1])]] <- res
}
```

Below we display a volcano plot for each comparison. These plots are also 
saved to pdf files. In each plot, a subset of the significant hits are 
indicated by name (n=`r volcanoMaxFeatures`, selected as the ones with 
the largest Manhattan distance to the origin). In addition, any features 
explicitly provided (see the [table above](#settings-table)) are also labeled. 
In addition to these pdf files, if "complexes" is specified to be included in 
the feature collections (and tested for significance using camera), we also 
generate a multi-page pdf file showing the position of the proteins of each 
significantly differentially abundant complex in the volcano plot. This pdf 
file is only generated if there is at least one significant complex (with 
adjusted p-value below the specified complexFDRThr=`r complexFDRThr`). 

```{r, echo=(stattest=="limma"), eval=(stattest=="limma"), warning=FALSE, fig.width=10.5, fig.height=7.5}
## Make a volcano plot for each comparison and save to a pdf file
xv <- "logFC"
yv <- "mlog10p"
apv <- "adj.P.Val"
volcind <- "showInVolcano"
for (nm in names(tests)) {
    xr <- range(tests[[nm]][[xv]], na.rm = TRUE)
    xr <- c(-max(abs(xr), na.rm = TRUE), max(abs(xr), na.rm = TRUE))
    gg0 <- ggplot(tests[[nm]], aes(x = .data[[xv]], y = .data[[yv]])) + 
        geom_point(fill = "lightgrey", color = "grey", pch = 21, size = 1.5) + 
        theme_bw() + coord_cartesian(xlim = xr) + 
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 14),
              title = element_text(size = 14)) + 
        annotate("text", x = min(xr), y = 0, hjust = "left",
                 vjust = "bottom", label = plotnotes[[nm]]) + 
        labs(x = "log2(fold change)", y = "-log10(p-value)", title = plottitles[[nm]],
             subtitle = paste0("Adj.p threshold = ", volcanoAdjPvalThr, 
                               ", |log2FC| threshold = ", volcanoLog2FCThr))
    gg <- gg0 + 
        geom_point(data = tests[[nm]] %>% 
                       dplyr::filter(.data[[volcind]]), 
                   fill = "red", color = "grey", pch = 21, size = 1.5) + 
        geom_text_repel(data = tests[[nm]] %>% 
                            dplyr::filter(.data[[volcind]] | 
                                              pid %in% volcanoFeaturesToLabel) %>%
                            dplyr::arrange(desc(abs(.data[[xv]]) + abs(.data[[yv]]))) %>%
                            dplyr::filter(between(row_number(), 0, volcanoMaxFeatures) | 
                                              pid %in% volcanoFeaturesToLabel),
                        aes(label = .data$pid), max.overlaps = Inf, size = 4,
                        min.segment.length = 0.1)
    print(gg)
    pdf(sub("\\.Rmd$", paste0("_volcano_", nm, ".pdf"), 
            knitr::current_input()), width = 10.5, height = 7.5)
    print(gg)
    dev.off()
    
    ## Create a volcano plot for each significantly enriched complex
    if ("complexes" %in% names(featureCollections)) {
        idx <- which(
            mcols(featureCollections$complexes)[, paste0(nm, "_FDR")] < complexFDRThr & 
                mcols(featureCollections$complexes)[, paste0(nm, "_NGenes")] > 1
        )
        tmpcomplx <- mcols(featureCollections$complexes)[idx, , drop = FALSE]
        tmpcomplx <- tmpcomplx[order(tmpcomplx[paste0(nm, "_PValue")]), , drop = FALSE]
        cplxs <- rownames(tmpcomplx)
        if (length(cplxs) > 0) {
            pdf(sub("\\.Rmd$", paste0("_volcano_", nm, "_complexes.pdf"), 
                    knitr::current_input()), width = 10.5, height = 7.5)
            for (cplx in cplxs) {
                prs <- featureCollections$complexes[[cplx]]
                cplxpval <- signif(mcols(featureCollections$complexes)[cplx, paste0(nm, "_PValue")], digits = 3)
                cplxfdr <- signif(mcols(featureCollections$complexes)[cplx, paste0(nm, "_FDR")], digits = 3)
                if (length(intersect(prs, tests[[nm]]$pid)) > 1) {
                    gg <- gg0 + 
                        geom_point(data = tests[[nm]] %>% 
                                       dplyr::filter(.data$pid %in% prs),
                                   fill = "red", color = "grey", pch = 21, size = 1.5) + 
                        geom_text_repel(data = tests[[nm]] %>% 
                                            dplyr::filter(.data$pid %in% prs),
                                        aes(label = .data$pid), max.overlaps = Inf, size = 4,
                                        min.segment.length = 0.1) + 
                        labs(caption = paste0(cplx, ", PValue = ", cplxpval,
                                              ", FDR = ", cplxfdr))
                    print(gg)
                    
                    ## Bar plot
                    bardata <- tests[[nm]] %>%
                        dplyr::filter(.data$pid %in% prs) %>%
                        dplyr::select(pid, matches("^Abundance")) %>%
                        tidyr::gather(key = "sample", value = "Abundance", -pid) %>%
                        dplyr::mutate(sample = sub("^Abundance.", "", sample)) %>%
                        dplyr::left_join(as.data.frame(colData(qft)), by = "sample") %>%
                        dplyr::filter(!is.na(group))
                    print(
                        ggplot(bardata %>% dplyr::group_by(pid, group) %>%
                                   dplyr::summarize(mean_Abundance = mean(Abundance, na.rm = TRUE),
                                                    sd_Abundance = sd(Abundance, na.rm = TRUE),
                                                    .groups = "drop"), 
                               aes(x = pid, y = mean_Abundance, fill = group)) +
                            geom_bar(position = position_dodge(), stat = "identity",
                                     colour = "black", size = 0.3) + 
                            geom_errorbar(aes(ymin = mean_Abundance - sd_Abundance,
                                              ymax = mean_Abundance + sd_Abundance), 
                                          size = 0.3, width = 0.2, 
                                          position = position_dodge(width = 0.9)) + 
                            geom_jitter(data = bardata, aes(y = Abundance), size = 2,
                                        position = position_dodge(width = 0.9)) + 
                            theme_bw() + 
                            theme(axis.text.x = element_text(size = 12, angle = 45, 
                                                             hjust = 0.5, 
                                                             vjust = 0.5),
                                  axis.text.y = element_text(size = 12), 
                                  axis.title = element_text(size = 14),
                                  title = element_text(size = 14)) + 
                            labs(x = "", y = "Mean +/- SD Abundance", title = cplx) + 
                            scale_fill_manual(name = "", 
                                              values = c("steelblue", "firebrick2"))
                    )
                }
            }
            dev.off()
        }
    }
}
```

```{r, echo=(stattest=="ttest"), eval=(stattest=="ttest"), warning=FALSE, fig.width=10.5, fig.height=7.5}
## Make a volcano plot for each comparison and save to a pdf file
xv <- "logFC"
yv <- "mlog10p"
apv <- "adj.P.Val"
tv <- "sam"
volcind <- "showInVolcano"
curvefun <- function(x, ta, s0, df) {
    -log10(2 * (1 - pt(q = ta * (1 + s0/(abs(x)/ta - s0)), df = df)))
}
for (nm in names(tests)) {
    xr <- range(tests[[nm]][[xv]], na.rm = TRUE)
    xr <- c(-max(abs(xr), na.rm = TRUE), max(abs(xr), na.rm = TRUE))
    yr <- range(tests[[nm]][[yv]], na.rm = TRUE)
    gg0 <- ggplot(tests[[nm]], aes(x = .data[[xv]], y = .data[[yv]])) + 
        geom_point(fill = "lightgrey", color = "grey", pch = 21, size = 1.5) + 
        theme_bw() + coord_cartesian(xlim = xr, ylim = yr) + 
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 14),
              title = element_text(size = 14)) + 
        annotate("text", x = min(xr), y = 0, hjust = "left",
                 vjust = "bottom", label = plotnotes[[nm]]) + 
        labs(x = "log2(fold change)", y = "-log10(p-value)", title = plottitles[[nm]],
             subtitle = paste0("FDR threshold = ", volcanoAdjPvalThr, 
                               ", s0 = ", curveparams[[nm]]$s0))
    if (is.finite(curveparams[[nm]]$ta)) {
        gg0 <- gg0 + 
            geom_line(data = data.frame(x = curveparams[[nm]]$x,
                                        y = curvefun(x = curveparams[[nm]]$x,
                                                     ta = curveparams[[nm]]$ta,
                                                     s0 = curveparams[[nm]]$s0,
                                                     df = curveparams[[nm]]$df)),
                      aes(x = x, y = y), color = "red", linetype = "dashed") + 
            geom_line(data = data.frame(x = -curveparams[[nm]]$x,
                                        y = curvefun(x = curveparams[[nm]]$x,
                                                     ta = curveparams[[nm]]$ta,
                                                     s0 = curveparams[[nm]]$s0,
                                                     df = curveparams[[nm]]$df)),
                      aes(x = x, y = y), color = "red", linetype = "dashed")
    }
    gg <- gg0 + 
        geom_point(data = tests[[nm]] %>% 
                       dplyr::filter(.data[[volcind]]),
                   fill = "red", color = "grey", pch = 21, size = 1.5) + 
        geom_text_repel(data = tests[[nm]] %>% 
                            dplyr::filter(.data[[volcind]] | 
                                              .data$pid %in% volcanoFeaturesToLabel) %>%
                            dplyr::arrange(desc(abs(.data[[xv]]) + abs(.data[[yv]]))) %>%
                            dplyr::filter(between(row_number(), 0, volcanoMaxFeatures) | 
                                              pid %in% volcanoFeaturesToLabel),
                        aes(label = .data$pid), max.overlaps = Inf, size = 4,
                        min.segment.length = 0.1)
    print(gg)
    pdf(sub("\\.Rmd$", paste0("_volcano_", nm, ".pdf"), 
            knitr::current_input()), width = 10.5, height = 7.5)
    print(gg)
    dev.off()
    
    ## Create a volcano plot for each significantly enriched complex
    if ("complexes" %in% names(featureCollections)) {
        idx <- which(
            mcols(featureCollections$complexes)[, paste0(nm, "_FDR")] < complexFDRThr & 
                mcols(featureCollections$complexes)[, paste0(nm, "_NGenes")] > 1
        )
        tmpcomplx <- mcols(featureCollections$complexes)[idx, , drop = FALSE]
        tmpcomplx <- tmpcomplx[order(tmpcomplx[paste0(nm, "_PValue")]), , drop = FALSE]
        cplxs <- rownames(tmpcomplx)
        if (length(cplxs) > 0) {
            pdf(sub("\\.Rmd$", paste0("_volcano_", nm, "_complexes.pdf"), 
                    knitr::current_input()), width = 10.5, height = 7.5)
            for (cplx in cplxs) {
                prs <- featureCollections$complexes[[cplx]]
                cplxpval <- signif(mcols(featureCollections$complexes)[cplx, paste0(nm, "_PValue")], digits = 3)
                cplxfdr <- signif(mcols(featureCollections$complexes)[cplx, paste0(nm, "_FDR")], digits = 3)
                if (length(intersect(prs, tests[[nm]]$pid)) > 1) {
                    gg <- gg0 + 
                        geom_point(data = tests[[nm]] %>% 
                                       dplyr::filter(.data$pid %in% prs),
                                   fill = "red", color = "grey", pch = 21, size = 1.5) + 
                        geom_text_repel(data = tests[[nm]] %>% 
                                            dplyr::filter(.data$pid %in% prs),
                                        aes(label = .data$pid), max.overlaps = Inf, size = 4,
                                        min.segment.length = 0.1) + 
                        labs(caption = paste0(cplx, ", PValue = ", cplxpval, 
                                              ", FDR = ", cplxfdr))
                    print(gg)
                    
                    ## Bar plot
                    bardata <- tests[[nm]] %>%
                        dplyr::filter(.data$pid %in% prs) %>%
                        dplyr::select(pid, matches("^Abundance")) %>%
                        tidyr::gather(key = "sample", value = "Abundance", -pid) %>%
                        dplyr::mutate(sample = sub("^Abundance.", "", sample)) %>%
                        dplyr::left_join(as.data.frame(colData(qft)), by = "sample") %>%
                        dplyr::filter(!is.na(group))
                    print(
                        ggplot(bardata %>% dplyr::group_by(pid, group) %>%
                                   dplyr::summarize(mean_Abundance = mean(Abundance, na.rm = TRUE),
                                                    sd_Abundance = sd(Abundance, na.rm = TRUE),
                                                    .groups = "drop"), 
                               aes(x = pid, y = mean_Abundance, fill = group)) +
                            geom_bar(position = position_dodge(), stat = "identity",
                                     colour = "black", size = 0.3) + 
                            geom_errorbar(aes(ymin = mean_Abundance - sd_Abundance,
                                              ymax = mean_Abundance + sd_Abundance), 
                                          size = 0.3, width = 0.2, 
                                          position = position_dodge(width = 0.9)) + 
                            geom_jitter(data = bardata, aes(y = Abundance), size = 2, 
                                        position = position_dodge(width = 0.9)) + 
                            theme_bw() + 
                            theme(axis.text.x = element_text(size = 12, angle = 45, 
                                                             hjust = 0.5, 
                                                             vjust = 0.5),
                                  axis.text.y = element_text(size = 12), 
                                  axis.title = element_text(size = 14),
                                  title = element_text(size = 14)) + 
                            labs(x = "", y = "Mean +/- SD Abundance", title = cplx) + 
                            scale_fill_manual(name = "", 
                                              values = c("steelblue", "firebrick2"))
                    )
                }
            }
            dev.off()
        }
    }
}
```

For each comparison, we also save a text file with the "significant" features 
(defined as those colored in the volcano plots above). The features are 
ordered by the logFC value. 

```{r}
for (nm in names(tests)) {
    write.table(tests[[nm]] %>% 
                    dplyr::filter(showInVolcano) %>%
                    dplyr::arrange(desc(logFC)), 
                file = sub("\\.Rmd$", paste0("_testres_", nm, ".txt"), 
                           knitr::current_input()),
                row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
}
```


```{r}
## Merge results from all tests
for (nm in names(tests)) {
    idx <- which(colnames(tests[[nm]]) != "pid")
    colnames(tests[[nm]])[idx] <- paste0(nm, ".", colnames(tests[[nm]])[idx])
}
all_tests <- as.data.frame(Reduce(function(...) dplyr::full_join(..., by = "pid"),
                                  tests), optional = TRUE)
rownames(all_tests) <- all_tests$pid
all_tests$pid <- NULL
```

# Assemble SingleCellExperiment object

Finally, we assemble all the information calculated above in a 
[SingleCellExperiment](https://bioconductor.org/packages/SingleCellExperiment/) 
object, which can later be used e.g. for exploration with 
[iSEE](https://bioconductor.org/packages/iSEE/).

```{r}
## Make 'base' SCE and add assays
sce <- qft[[aName]]
assayNames(sce) <- aName
for (nm in names(qft)) {
    assay(sce, nm) <- assay(qft[[nm]])
}

## Add sample and feature annotations
colData(sce) <- colData(qft)
stopifnot(rownames(sce) == rownames(all_tests))
rowData(sce) <- cbind(rowData(sce), all_tests)

## TODO
## Move some values to assays rather than rowData
colnames(rowData(sce)) <- gsub("\\.+$", "", colnames(rowData(sce)))
colnames(rowData(sce)) <- gsub("\\.+", ".", colnames(rowData(sce)))
moveToAssay <- c("Abundances.Grouped.", "Abundances.Grouped.CV.in.Percent.",
                 "Abundances.Grouped.Count.",
                 "Abundances.Normalized.", "Found.in.Sample.Group.")
for (mta in moveToAssay) {
    cols <- sub(iColPattern, mta, colnames(sce))
    if (all(cols %in% colnames(rowData(sce)))) {
        ## Use withDimnames = FALSE since colnames are not identical
        assay(sce, sub("\\.$", "", mta), withDimnames = FALSE) <- 
            as.matrix(rowData(sce)[, cols])
        ## For iBAQ, also include log-transformed values (for use in heatmaps)
        # if (mta == "iBAQ." && !("log2_iBAQ_withNA" %in% assayNames(sce))) {
        #     tmplogibaq <- log2(as.matrix(rowData(sce)[, cols]))
        #     tmplogibaq[!is.finite(tmplogibaq)] <- NA
        #     assay(sce, paste0("log2_", sub("\\.$", "", mta), "_withNA"), 
        #           withDimnames = FALSE) <- tmplogibaq
        # }
    }
    ## Remove all columns corresponding to this assay from the rowData, 
    ## even if only some samples are retained
    colsall <- sub(iColPattern, mta, iColsAll)
    rowData(sce) <- rowData(sce)[, !colnames(rowData(sce)) %in% colsall]
}

## TODO
## Remove some columns from rowData (save to text file)
colsToRemove <- c(grep("Found.in.Fraction", colnames(rowData(sce)), value = TRUE))
write.table(as.data.frame(rowData(sce)[, colsToRemove]) %>%
                tibble::rownames_to_column("ID"), 
            file = sub("\\.Rmd$", paste0("_sce_extra_annots.tsv"), 
                       knitr::current_input(dir = TRUE)), 
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
rowData(sce) <- rowData(sce)[, !colnames(rowData(sce)) %in% colsToRemove]

## Add information about missing values
nacols <- as.data.frame(nbr_na$nNAcols) %>%
    dplyr::filter(assay == aName)
stopifnot(all(sub(iColPattern, "", nacols$name) == sce$sample))
colData(sce) <- cbind(colData(sce), nacols[, c("nNA", "pNA")])
narows <- as.data.frame(nbr_na$nNArows) %>%
    dplyr::filter(assay == aName)
stopifnot(all(narows$name == rownames(sce)))
rowData(sce) <- cbind(rowData(sce), narows[, c("nNA", "pNA")])

## Add experiment metadata
metadata(sce) <- c(metadata(sce), 
                   list(
                       pdFile = pdFile,
                       aName = aName,
                       iColPattern = iColPattern,
                       imputeMethod = imputeMethod,
                       ctrlGroup = ctrlGroup,
                       allPairwiseComparisons = allPairwiseComparisons,
                       normMethod = normMethod,
                       stattest = stattest,
                       minlFC = minlFC,
                       analysisDate = as.character(Sys.Date()),
                       rmdFile = knitr::current_input(dir = TRUE)
                   ))

## Register logFC/AveAb/pvalue fields for use in iSEE
sce <- iSEEu::registerLogFCFields(
    sce, grep("\\.logFC$", colnames(rowData(sce)), value = TRUE)
)
sce <- iSEEu::registerAveAbFields(
    sce, grep("\\.AveExpr$", colnames(rowData(sce)), value = TRUE)
)
sce <- iSEEu::registerPValueFields(
    sce, grep("\\.P.Value$", colnames(rowData(sce)), value = TRUE)
)

sce <- registerFeatureSetCollections(sce, featureCollections)

sce
```


# Run PCA

We run a principal component analysis to obtain a reduced dimensionality 
representation of the data, in order to visualize the samples in two dimensions. 

```{r}
sce <- as(sce, "SingleCellExperiment")
sce <- scater::runPCA(sce, exprs_values = assayForTests, name = "PCA", 
                      ncomponents = 10, ntop = Inf,
                      BSPARAM = BiocSingular::ExactParam())
scater::plotReducedDim(sce, dimred = "PCA", colour_by = "group",
                       point_size = 5)
```

# Heatmap with hierarchical clustering

For another birds-eye view of the data, we represent it using a heatmap of 
the (imputed and normalized) log intensities, and cluster the samples and 
proteins using hierarchical clustering. In the first heatmap below, the values 
represent the normalized log intensities directly. In the second heatmap, 
the values for each protein have been centered to mean 0.

```{r, fig.height=8}
# gg_color_hue <- function(n) {
#     hues = seq(15, 375, length = n + 1)
#     hcl(h = hues, l = 65, c = 100)[1:n]
# }
# groupcols = gg_color_hue(length(unique(sce$group)))
# names(groupcols) <- levels(factor(sce$group))
# 
# ht <- ComplexHeatmap::Heatmap(
#     assay(sce, assayForTests),
#     name = assayForTests,
#     show_row_names = FALSE,
#     use_raster = TRUE,
#     top_annotation = columnAnnotation(group = sce$group,
#                                       col = list(group = groupcols))
# )
# draw(ht, merge_legends = TRUE)
# 
# ## Centered
# ht <- ComplexHeatmap::Heatmap(
#     t(scale(t(assay(sce, assayForTests)), center = TRUE, scale = FALSE)),
#     name = paste0(assayForTests, "\ncentered"),
#     show_row_names = FALSE,
#     use_raster = TRUE,
#     top_annotation = columnAnnotation(group = sce$group,
#                                       col = list(group = groupcols))
# )
# draw(ht, merge_legends = TRUE)
```


# Save SingleCellExperiment object

The `SingleCellExperiment` object created above is saved in the following location: 

```{r}
sceFile <- sub("\\.Rmd$", paste0("_sce.rds"), knitr::current_input(dir = TRUE))
saveRDS(sce, file = sceFile)
sceFile
```

In addition, all feature information (the `rowData` of the `SingleCellExperiment`)
is written to a text file:

```{r}
textFile <- sub("\\.Rmd$", paste0("_feature_info.txt"), 
                knitr::current_input(dir = TRUE))
write.table(as.data.frame(rowData(sce)) %>% 
                tibble::rownames_to_column("FeatureID"), 
            file = textFile, row.names = FALSE, col.names = TRUE,
            quote = FALSE, sep = "\t")
textFile
```


# Explore the data interactively

To explore the data further using [iSEE](https://bioconductor.org/packages/iSEE/), 
navigate to http://rstudio.fmi.ch and type the following line of code into 
the R console: 

```{r, echo=FALSE}
iSEEScript <- sub("\\.Rmd$", paste0("_iSEE.R"), knitr::current_input(dir = TRUE))
```


```{r create-markdown-chunks-dynamically, include=FALSE}
out <- paste0("\n```{r, results='asis', echo=TRUE, eval=FALSE, class.source = 'fold-show'}\n\nsource('", 
              iSEEScript, "')\n\n```")
```

`r paste(knitr::knit(text = out), collapse = '\n')`


That will open up an iSEE session where you can interactively explore your data.

```{r, echo=FALSE}
## Assemble a script that can be sourced to run iSEE
iSEECode <- c(
    "library(iSEE)",
    "library(iSEEu)",
    "library(shiny)",
    paste0("sce <- readRDS('", sceFile, "')"),
    "imp_color_fun <- function(n) {",
    "    structure(c('grey', 'firebrick1'), names = c('TRUE', 'FALSE'))",
    "}",
    "ecm <- ExperimentColorMap(",
    "    assays = list(",
    paste0("        imputed_", aName, " = imp_color_fun"),
    "    )",
    ")",
    "app <- iSEE(sce, colormap = ecm, initial = list("
)
for (nm in names(tests)) {
    iSEECode <- c(
        iSEECode,
        c(paste0("    VolcanoPlot(XAxis = 'Row data', YAxis = '", nm, ".P.Value',"),
          paste0("                XAxisRowData = '", nm, ".logFC', "),
          "                RowSelectionDynamicSource = TRUE, PanelWidth = 4L),")
    )
}
for (nm in names(tests)) {
    iSEECode <- c(
        iSEECode,
        c(paste0("    MAPlot(XAxis = 'Row data', YAxis = '", nm, ".logFC',"),
          paste0("           XAxisRowData = '", nm, ".AveExpr', "),
          paste0("           PValueField = '", nm, ".P.Value', "),
          "           RowSelectionDynamicSource = TRUE, PanelWidth = 4L),")
    )
}
iSEECode <- c(
    iSEECode, 
    "    RowDataTable(PanelWidth = 6L, RowSelectionDynamicSource = TRUE),",
    "    FeatureAssayPlot(PanelWidth = 3L, ", 
    paste0("                     Assay = '", assayForTests, "',"),
    "                     XAxis = 'Column data', XAxisColumnData = 'group',",
    "                     ColorBy = 'Feature name',", 
    paste0("                     ColorByFeatureNameAssay = 'imputed_", aName, "',"),
    "                     ColorByFeatureSource = 'RowDataTable1',",
    "                     YAxisFeatureSource = 'RowDataTable1', PointSize = 5),",
    "    ReducedDimensionPlot(PanelWidth = 3L, PointSize = 5, ColorBy = 'Column data',",
    "                         ColorByColumnData = 'group'),",
    "    FeatureSetTable(PanelWidth = 8L), ",
    "    ComplexHeatmapPlot(PanelWidth = 4L, ",
    paste0("                       Assay = 'Abundance', "),
    "                       RowSelectionDynamicSource = TRUE, CustomRows = FALSE,",
    "                       ColumnData = 'group', ShowColumnSelection = FALSE,", 
    "                       OrderColumnSelection = FALSE), ",
    "    SampleAssayPlot(PanelWidth = 4L, ", 
    paste0("                    Assay = '", assayForTests, "',"),
    "                    XAxis = 'Sample name'),",
    "    ColumnDataPlot(PanelWidth = 4L, YAxis = 'pNA', XAxis = 'Column data',",
    "                   XAxisColumnData = 'group', ColorBy = 'Column data',",
    "                   PointSize = 5, ColorByColumnData = 'group'),",
    "    RowDataPlot(PanelWidth = 4L, YAxis = 'Score')",
    "))",
    "shiny::runApp(app)"
)

writeLines(iSEECode, con = iSEEScript, sep = "\n")
```

# Session info {#session-info}

This report was compiled with the following package versions:

```{r, echo=FALSE}
sessionInfo()
```

# References


