{{YmlParametersStart}}

{{YmlParametersEnd}}

---
subtitle: "`r paste0('Generated with einprot v', packageVersion('einprot'))`"
date: "`r Sys.time()`"
output: 
    html_document:
        theme: united
        toc: true
        toc_float: true
        code_folding: hide
editor_options: 
  chunk_output_type: console
references:
- id: Ritchie2015limma
  title: limma powers differential expression analyses for RNA-sequencing and microarray studies
  author:
  - family: Ritchie
    given: M E
  - family: Phipson
    given: B
  - family: Wu
    given: D
  - family: Hu
    given: Y
  - family: Law
    given: C W
  - family: Shi
    given: W
  - family: Smyth
    given: G K
  container-title: Nucleic Acids Research
  volume: 43
  issue: 7
  page: e47
  type: article-journal
  URL: https://academic.oup.com/nar/article/43/7/e47/2414268
  issued:
    year: 2015
- id: McCarthy2009treat
  title: Testing significance relative to a fold-change threshold is a TREAT
  author:
  - family: McCarthy
    given: D J
  - family: Smyth
    given: G K
  container-title: Bioinformatics
  volume: 25
  page: 765-771
  type: article-journal
  URL: http://bioinformatics.oxfordjournals.org/content/25/6/765
  issued:
    year: 2009
- id: Wu2012camera
  title: Camera:a competitive gene set test accounting for inter-gene correlation
  author:
  - family: Wu
    given: D
  - family: Smyth
    given: G K
  container-title: Nucleic Acids Research
  volume: 40
  issue: 17
  page: e133
  type: article-journal
  URL: https://academic.oup.com/nar/article/40/17/e133/2411151
  issued:
    year: 2012
- id: Phipson2016robust
  title: Robust hyperparameter estimation protects against hypervariable genes and improves power to detect differential expression
  author:
  - family: Phipson
    given: B
  - family: Lee
    given: S
  - family: Majewski
    given: I J
  - family: Alexander
    given: W S
  - family: Smyth
    given: G K
  container-title: Annals of Applied Statistics 
  volume: 10
  page: 946-963
  type: article-journal
  URL: http://projecteuclid.org/euclid.aoas/1469199900
  issued:
    year: 2016
- id: Tusher2001sam
  title: Significance analysis of microarrays applied to the ionizing radiation response
  author:
  - family: Tusher
    given: V G
  - family: Tibshirani
    given: R
  - family: Chu
    given: G
  container-title: Proceedings of the National Academy of Sciences of the United States of America 
  volume: 98
  page: 5116-5121
  type: article-journal
  URL: https://www.pnas.org/content/98/9/5116
  issued:
    year: 2001
- id: Tyanova2016perseus
  title: The Perseus computational platform for comprehensive analysis of (prote)omics data
  author:
  - family: Tyanova
    given: S
  - family: Temu
    given: T
  - family: Sinitcyn
    given: P
  - family: Carlson
    given: A
  - family: Hein
    given: M Y
  - family: Geiger
    given: T
  - family: Mann
    given: M
  - family: Cox
    given: J
  container-title: Nature Methods 
  volume: 13
  page: 731-740
  type: article-journal
  URL: https://www.nature.com/articles/nmeth.3901
  issued:
    year: 2016
- id: Cox2008maxquant
  title: MaxQuant enables high peptide identification rates, individualized p.p.b.-range mass accuracies and proteome-wide protein quantification
  author:
  - family: Cox
    given: J
  - family: Mann
    given: M
  container-title: Nature Biotechnology
  volume: 26
  page: 1367-1372
  type: article-journal
  URL: https://www.nature.com/articles/nbt.1511
  issued:
    year: 2008
- id: Szklarczyk2021string
  title: The STRING database in 2021 - customizable protein-protein networks, and functional characterization of user-uploaded gene/measurement sets
  author:
  - family: Szklarczyk 
    given: D
  - family: Gable
    given: AL
  - family: Nastou
    given: KC
  - family: Lyon
    given: D
  - family: Kirsch
    given: R
  - family: Pyysalo
    given: S
  - family: Doncheva
    given: NT
  - family: Legeay
    given: M
  - family: Fang
    given: T
  - family: Bork
    given: P
  - family: Jensen
    given: LJ
  - family: von Mering
    given: C
  container-title: Nucleic Acids Research
  volume: 49(D1)
  page: D605-D612
  type: article-journal
  URL: https://academic.oup.com/nar/article/49/D1/D605/6006194
  issued:
    year: 2021
---

{{ConfigParametersStart}}

{{ConfigParametersEnd}}

---
title: "`r reportTitle`"
author: "`r reportAuthor`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 200)
```

```{r load-pkg, echo=FALSE}
## Load packages
suppressPackageStartupMessages({
    library(QFeatures)
    library(SummarizedExperiment)
    library(SingleCellExperiment)
    library(imputeLCMD)
    library(rrcovNA)
    library(scater)
    library(dplyr)
    library(tibble)
    library(tidyr)
    library(ggplot2)
    library(ggrepel)
    library(cowplot)
    library(plotly)
    library(UpSetR)
    library(circlize)
    library(ComplexHeatmap)
    library(DT)
    library(STRINGdb)
    library(MsCoreUtils)
    library(einprot, lib.loc = "/tungstenfs/groups/gbioinfo/sonechar/FMIgroups/Proteomics/devel_lib_einprot/")
})
```

This report describes a reproducible end-to-end analysis of a proteomics dataset quantified 
with [MaxQuant](https://www.maxquant.org/) [@Cox2008maxquant]. 
Most of the code is hidden by default, 
but can be displayed by clicking on the `Code` buttons (or by selecting 
`Code -> Show All Code` in the top right corner of the report).

```{r get-basic-info}
## Get species info and define STRINGdb object
speciesInfo <- getSpeciesInfo(species)
string_db <- STRINGdb$new(version = "11.5", species = speciesInfo$taxId, 
                          score_threshold = 400, input_directory = "")

## If needed and not provided, define path to complex DB 
## (will be added to summary table below)
if ("complexes" %in% includeFeatureCollections && is.null(complexDbPath)) {
    complexDbPath <- system.file(EINPROT_COMPLEXES_FILE,
                                 package = "einprot", 
                                 lib.loc = "/tungstenfs/groups/gbioinfo/sonechar/FMIgroups/Proteomics/devel_lib_einprot/")
}

## Get conversion tables for PomBase and WormBase IDs
pbconv <- readRDS(system.file(EINPROT_POMBASE_CONVTABLE,
                              package = "einprot", 
                              lib.loc = "/tungstenfs/groups/gbioinfo/sonechar/FMIgroups/Proteomics/devel_lib_einprot/"))
wbconv <- readRDS(system.file(EINPROT_WORMBASE_CONVTABLE,
                              package = "einprot", 
                              lib.loc = "/tungstenfs/groups/gbioinfo/sonechar/FMIgroups/Proteomics/devel_lib_einprot/"))
```

# Experiment details

```{r exp-table}
makeTableFromList(c(experimentInfo, 
                    list(
                        "Species" = speciesInfo$species,
                        "Species (common)" = speciesInfo$speciesCommon,
                        "Taxonomic ID" = speciesInfo$taxId
                    )))
```

# MaxQuant analysis summary

```{r mq-table}
mq <- readMaxQuantXML(mqParameterFile)
mq <- c(list("MaxQuant file" = mqFile), mq)
makeTableFromList(mq)
```

# Settings {#settings-table}

```{r settings-table}
settingsList <- list(
    "Column pattern" = iColPattern,
    "Include only samples (if applicable)" = paste(includeOnlySamples, 
                                                   collapse = ", "),
    "Exclude samples" = paste(excludeSamples, collapse = ", "),
    "Min. number of peptides" = minPeptides,
    "Min. protein score" = minScore,
    "Imputation method" = imputeMethod,
    "Min. nbr valid values" = minNbrValidValues,
    "Model fit" = ifelse(singleFit, "Single (one model fit for all samples)", 
                         "Separate model fit for each comparison"),
    "Groups to merge" = paste(unlist(
        lapply(names(mergeGroups), 
               function(nm) paste0(nm, ":", paste(mergeGroups[[nm]], collapse = ",")))),
        collapse = "; "),
    "Comparisons" = paste(unlist(lapply(comparisons, 
                                        function(x) paste(x, collapse = " vs "))),
                          collapse = "; "),
    "Control group" = ctrlGroup,
    "Do all pairwise comparisons" = allPairwiseComparisons,
    "Normalization method" = normMethod,
    "Statistical test" = stattest,
    "Minimal fold change (limma/treat)" = minlFC,
    "Adjusted p-value threshold for volcano plots" = volcanoAdjPvalThr,
    "Log2 FC threshold for volcano plots" = volcanoLog2FCThr,
    "Max nbr features to indicate in volcano plots" = volcanoMaxFeatures,
    "s0" = volcanoS0,
    "Features to always label in volcano plots" = paste(volcanoFeaturesToLabel,
                                                        collapse = ", "),
    "Feature collections" = paste(includeFeatureCollections, collapse = "; "),
    "Min size to keep feature set" = minSizeToKeepSet,
    "Complexes file" = gsub(".+\\/(.+.rds)", "\\1", complexDbPath),
    "Complexes from species" = complexSpecies,
    "Custom complexes" = names(customComplexes),
    "FDR Threshold for complexes" = complexFDRThr,
    "Max nbr complexes to plot" = maxNbrComplexesToPlot,
    "Number of permutations" = nperm,
    "Random seed" = seed)

if (stattest == "ttest") {
    settingsList <- settingsList[!(names(settingsList) %in% 
                                       c("Minimal fold change (limma/treat)",
                                         "Log2 FC threshold for volcano plots"))]
}
if (stattest == "limma") {
    settingsList <- settingsList[!(names(settingsList) %in% 
                                       c("s0", "Number of permutations"))]
}

makeTableFromList(settingsList)
```


# Read MaxQuant output

The input to this workflow is a `proteinGroups.txt` file from MaxQuant 
(see path in the table above). We read the MaxQuant intensities into `R` 
and store them in a 
[SingleCellExperiment](https://bioconductor.org/packages/SingleCellExperiment/) 
object. This object will later be expanded with additional data, such as 
transformed and imputed abundances. 

At this point, the SingleCellExperiment object holds assays with the different 
types of intensities and annotations from the MaxQuant file.

```{r read-data}
## Read MaxQuant output
tmp <- importExperiment(inFile = mqFile, iColPattern = iColPattern,
                        includeOnlySamples = includeOnlySamples,
                        excludeSamples = excludeSamples)
sce <- tmp$sce
aName <- tmp$aName

## Make sure that the columns used for filtering later are character vectors
rowData(sce)$Reverse[is.na(rowData(sce)$Reverse)] <- ""
rowData(sce)$Potential.contaminant[is.na(rowData(sce)$Potential.contaminant)] <- ""
rowData(sce)$Only.identified.by.site[is.na(rowData(sce)$Only.identified.by.site)] <- ""

sce
```

# Modify feature names

The feature ID used when reading the data above are numeric indices. 
We replace these IDs with more interpretable ones, corresponding to 
the first annotated gene name, or, in cases where that is missing, with 
the first annotated majority protein ID. When the gene name doesn't uniquely 
identify the feature, it is combined with the majority protein ID to generate 
the final feature ID. 

```{r fix-ids}
sce <- fixFeatureIds(sce, geneIdCol = "Gene.names",
                     proteinIdCol = "Majority.protein.IDs")
```

# Add sample annotations

Next, we compile the sample annotations. The sample IDs have been extracted 
from the column names in the `MaxQuant` file, by removing the provided 
`iColPattern` from the main intensity columns. The `group` 
column will be used to define groups for the statistical testing later. 
Please check that the table below correspond to your expectations.

```{r add-sampleinfo}
sce <- addSampleAnnots(sce, sampleAnnot = sampleAnnot, mergeGroups = mergeGroups)

DT::datatable(as.data.frame(colData(sce)),
              options = list(scrollX = TRUE, pageLength = 20))
```

# Overall distribution of protein intensities

The box plot below displays the distribution of raw protein intensities in 
each sample, on a log scale (excluding any missing values).

```{r intensity-distribution, warning=FALSE, message=FALSE}
makeIntensityBoxplots(sce = sce, assayName = aName, doLog = TRUE, 
                      ylab = "Raw intensity")
```

# Filter out contaminants, reverse hits, and proteins with low confidence

Next, we filter out any features classified by MaxQuant as potential 
contaminants or reverse (decoy) hits, and features identified only by site (which 
removes proteins that are only identified by peptides carrying one or more 
modified amino acids). In addition, we may remove protein identifications 
based on score and number of peptides (i.e. to exclude one-hit wonders). 
The UpSet plot below illustrates the overlaps between the sets of features 
filtered out based on the different criteria (vertical bars). 

```{r filter-features}
nbrFeaturesBefore <- nrow(sce)
filtdf <- as.data.frame(rowData(sce)) %>%
    dplyr::select(Reverse, Potential.contaminant, Only.identified.by.site,
                  Score, Peptides) %>%
    dplyr::mutate(across(c(Reverse, Potential.contaminant, 
                           Only.identified.by.site), 
                         function(x) as.numeric(x == "+"))) %>%
    dplyr::mutate(Score = as.numeric((Score < minScore)|is.na(Score)),
                  Peptides = as.numeric((Peptides < minPeptides)|is.na(Peptides)))
sce <- sce[which(rowData(sce)$Reverse == "" & 
                   rowData(sce)$Potential.contaminant == "" & 
                   rowData(sce)$Only.identified.by.site == "" & 
                   rowData(sce)$Score >= minScore & 
                   rowData(sce)$Peptides >= minPeptides), ]
nbrFeaturesAfter <- nrow(sce)

UpSetR::upset(filtdf)
```

This filtering removed `r nbrFeaturesBefore - nbrFeaturesAfter` features. 
The new `sce` object has `r nbrFeaturesAfter` features.

# Prepare feature collections for later testing

In addition to testing individual proteins for differential abundance 
between groups, we can also test collections of proteins. Here we 
define the collections that will be used. 

```{r prepare-feature-collections}
(featureCollections <- prepareFeatureCollections(
    sce = sce, idCol = "Gene.names", 
    includeFeatureCollections = includeFeatureCollections,
    complexDbPath = complexDbPath, speciesInfo = speciesInfo,
    complexSpecies = complexSpecies, customComplexes = customComplexes,
    minSizeToKeep = minSizeToKeepSet))
```

# Apply a log2 transformation

Before the downstream analysis, we log2-transform the measured intensities. We also 
add an additional assay to keep track of the position of the missing values
(which will be imputed later). 

```{r log-transform}
assay(sce, paste0("log2_", aName, "_withNA")) <- log2(assay(sce, aName))

## Add assay indicating missing values, which will be imputed
tmp <- assay(sce, paste0("log2_", aName, "_withNA"))
tmp <- !is.finite(tmp)
assay(sce, paste0("imputed_", aName)) <- tmp
sce

## Set the assay to use for tests later (depending on the downstream steps,
## this may be updated below)
assayForTests <- paste0("log2_", aName)
```

# Visualize missing value patterns

The plot below shows the fraction of the total set of features that are detected 
(with a non-missing value) in each of the samples (based on the `r aName` assay).

```{r missing-values}
## Replace zeros/-Inf values by explicit NA values in the assays
assay(sce, aName)[assay(sce, aName) == 0] <- NA
assay(sce, paste0("log2_", aName, "_withNA"))[
  !is.finite(assay(sce, paste0("log2_", aName, "_withNA")))] <- NA

## Count number of NA values and add to SCE
nbr_na <- nNA(sce)
nbr_na <- lapply(nbr_na, function(a) {
    a$assay <- assayNames(sce)[1]
    a
})
nacols <- as.data.frame(nbr_na$nNAcols)
stopifnot(all(nacols$name == sce$sample))
SummarizedExperiment::colData(sce) <-
    cbind(SummarizedExperiment::colData(sce), nacols[, c("nNA", "pNA")])
narows <- as.data.frame(nbr_na$nNArows)
stopifnot(all(narows$name == rownames(sce)))
SummarizedExperiment::rowData(sce) <-
    cbind(SummarizedExperiment::rowData(sce), narows[, c("nNA", "pNA")])

plotFractionDetectedPerSample(dfNA = nbr_na$nNAcols, aName = aName)
```

We also plot the number of features that are detected in a given number 
of samples.

```{r missing-values-2}
plotDetectedInSamples(dfNA = nbr_na$nNArows, aName = aName)
```

Finally we show the full missing value structure in the data matrix. The white 
values correspond to missing values, and the grey ones to observed intensities.

```{r missing-values-overall, fig.height=7, message=FALSE}
plotMissingValuesHeatmap(sce, assayMissing = paste0("imputed_", aName))
```


# Imputation

Next, we apply the `r imputeMethod` method to perform imputation 
of the log2-transformed data, and plot the distribution of imputed and 
non-imputed values in each sample.


```{r plot-imputation, fig.width=7, fig.height=7, message = FALSE}
set.seed(seed)
sce <- doImputation(sce, method = imputeMethod, 
                    assayName = paste0("log2_", aName, "_withNA"),
                    imputedAssayName = paste0("log2_", aName))
plotImputationDistribution(sce, assayToPlot = paste0("log2_", aName), 
                           assayImputation = paste0("imputed_", aName),
                           xlab = "log2 intensity")
```

# Overall distribution of log2 protein intensities

Next we consider the overall distribution of log2-intensities among the 
samples (after imputation).

```{r intensity-distribution-imputed}
makeIntensityBoxplots(sce = sce, assayName = paste0("log2_", aName),
                      doLog = FALSE, 
                      ylab = "log2 intensity (after imputation)")
```

```{r text-norm, results="asis", echo=FALSE}
cat(normText(normMethod = normMethod))
```

```{r normalize, echo=(normMethod != "none"), eval=(normMethod != "none"), message=FALSE}
assay(sce, paste0("log2norm_", aName)) <- 
    normalize_matrix(assay(sce, paste0("log2_", aName)),
                     method = normMethod)

## If data is normalized, use normalized values for testing
assayForTests <- paste0("log2norm_", aName)

makeIntensityBoxplots(sce = sce, assayName = paste0("log2norm_", aName),
                      doLog = FALSE, 
                      ylab = "log2 intensity (after imputation and normalization)")
```

```{r remove-batch-effect}
assayForPlots <- assayForTests
if ("batch" %in% colnames(colData(sce))) {
    assayForPlots <- paste0(assayForTests, "_batchcorr")
    assay(sce, assayForPlots) <- 
        limma::removeBatchEffect(assay(sce, assayForTests), batch = sce$batch, 
                                 design = model.matrix(~ sce$group))
}
```

# Statistical testing

For each feature, we then compare the (possibly imputed) log2 intensities 
between groups. 

```{r text-test, results="asis", echo=FALSE}
cat(testText(testType = stattest))
```

The `r assayForTests` assay will be used for the tests. The 
following pairwise comparisons will be performed (in each case, the 
first listed group will be the 'baseline' group): 

```{r initialize-tests}
(comparisons <- makeListOfComparisons(
    allGroups = unique(sce$group), comparisons = comparisons, 
    allPairwiseComparisons = allPairwiseComparisons,
    ctrlGroup = ctrlGroup))
```


```{r run-test}
aval <- ifelse("iBAQ" %in% assayNames(sce), "iBAQ", aName)
testres <- runTest(sce = sce, comparisons = comparisons, testType = stattest, 
                   assayForTests = assayForTests,
                   assayImputation = paste0("imputed_", aName), 
                   minNbrValidValues = minNbrValidValues,
                   minlFC = minlFC, featureCollections = featureCollections, 
                   complexFDRThr = complexFDRThr,
                   volcanoAdjPvalThr = volcanoAdjPvalThr, 
                   volcanoLog2FCThr = volcanoLog2FCThr,
                   baseFileName = sub("\\.Rmd$", "", knitr::current_input()),
                   seed = seed, nperm = nperm, volcanoS0 = volcanoS0, 
                   aName = "iBAQ", addAbundanceValues = TRUE, 
                   singleFit = singleFit)
tests <- testres$tests
featureCollections <- testres$featureCollections

for (cmp in comparisons) {
    ## Add WormBase/PomBase IDs if applicable
    if (speciesInfo$speciesCommon == "roundworm") {
        tests[[paste0(cmp[2], "_vs_", cmp[1])]]$WormBaseID <- 
            wbconv$WormBaseID[match(tests[[paste0(cmp[2], "_vs_", cmp[1])]]$proteinIdSingle, wbconv$UniProtID)]
    } else if (speciesInfo$speciesCommon == "fission yeast") {
        tests[[paste0(cmp[2], "_vs_", cmp[1])]]$PomBaseID <- 
            pbconv$PomBaseID[match(tests[[paste0(cmp[2], "_vs_", cmp[1])]]$proteinIdSingle, pbconv$UniProtID)]
    }
}
```

```{r text-expdesign, results="asis", echo=FALSE}
cat(expDesignText(testType = stattest))
```

```{r expdesign-plot, eval = (stattest == "limma"), echo = (stattest == "limma"), fig.width = 12, fig.height = 8}
if ("design" %in% names(testres$design)) {
    title <- cowplot::ggdraw() + cowplot::draw_label("Overall design", 
                                                     x = 0, hjust = 0,
                                                     fontface = "bold")
    vd <- ExploreModelMatrix::VisualizeDesign(
        testres$design$sampleData, designFormula = NULL, 
        designMatrix = testres$design$design)
    print(cowplot::plot_grid(
        title, cowplot::plot_grid(
            plotlist = c(vd$plotlist, vd$cooccurrenceplots), nrow = 1), 
        ncol = 1, rel_heights = c(0.08, 1))
    )
    print(testres$design$design)
    print(testres$design$contrasts)
} else {
    for (nm in names(testres$design)) {
        title <- cowplot::ggdraw() + cowplot::draw_label(nm, x = 0, hjust = 0)
        vd <- ExploreModelMatrix::VisualizeDesign(
            testres$design[[nm]]$sampleData, designFormula = NULL, 
            designMatrix = testres$design[[nm]]$design)
        print(cowplot::plot_grid(
            title, cowplot::plot_grid(
                plotlist = c(vd$plotlist, vd$cooccurrenceplots), nrow = 1),
            ncol = 1, rel_heights = c(0.08, 1))
        )
        print(testres$design[[nm]]$design)
        print(testres$design[[nm]]$contrast)
    }
}
```


```{r test-messages, echo = (length(testres$messages) > 0), eval = (length(testres$messages) > 0)}
testres$messages
```

```{r text-sa, results="asis", echo=FALSE}
cat(saText(testType = stattest))
```
 
```{r plot-sa, fig.width = 12, fig.height = ceiling(length(tests)/3) * 4, warning = FALSE, eval = (stattest == "limma"), echo = (stattest == "limma")}
xrng <- range(unlist(lapply(tests, function(df) df$AveExpr)), na.rm = TRUE)
yrng <- range(unlist(lapply(tests, function(df) sqrt(df$sigma))), na.rm = TRUE)
saplots <- lapply(names(tests), function(nm) {
    df <- tests[[nm]]
    ggplot(df) + 
        geom_point(aes(x = AveExpr, y = sqrt(sigma)), alpha = 0.5, size = 0.5) + 
        geom_line(aes(x = AveExpr, y = sqrt(sqrt(s2.prior))), color = "blue", 
                  size = 1) + 
        labs(x = "Average log abundance", 
             y = "sqrt(residual standard deviation)",
             title = nm) + 
        coord_cartesian(xlim = xrng, ylim = yrng) + 
        theme_bw()
})
cowplot::plot_grid(plotlist = saplots, ncol = 3)
```


Below we display a volcano plot for each comparison. These plots are also 
saved to pdf files. In each plot, a subset of the significant hits are 
indicated by name (n=`r volcanoMaxFeatures`, selected as the ones with 
the largest Manhattan distance to the origin). These proteins are also used 
to generate STRING networks [@Szklarczyk2021string] (separately for the 
up- and downregulated ones), which are included in the pdf file. Any features 
explicitly requested (see the [table above](#settings-table)) are also labeled
in the volcano plots. 
In addition to these pdf files, if "complexes" is specified to be included in 
the feature collections (and tested for significance using camera), we also 
generate a multi-page pdf file showing the position of the proteins of each 
significantly differentially abundant complex in the volcano plot, as well 
as bar plots of the proteins' iBAQ values in the compared samples. This pdf 
file is only generated if there is at least one significant complex (with 
adjusted p-value below the specified complexFDRThr=`r complexFDRThr`). 

```{r volcano-limma, echo=(stattest=="limma"), eval=(stattest=="limma"), warning=FALSE, fig.width=10.5, fig.height=7.5}
## Make a volcano plot for each comparison and save to a pdf file
xv <- "logFC"
yv <- "mlog10p"
xvma = "AveExpr"
apv <- "adj.P.Val"
volcind <- "showInVolcano"
interactiveVolcanos <- htmltools::tagList()
```

```{r volcano-ttest, echo=(stattest=="ttest"), eval=(stattest=="ttest"), warning=FALSE, fig.width=10.5, fig.height=7.5}
## Make a volcano plot for each comparison and save to a pdf file
xv <- "logFC"
yv <- "mlog10p"
xvma = NULL
apv <- "adj.P.Val"
tv <- "sam"
volcind <- "showInVolcano"
interactiveVolcanos <- htmltools::tagList()
```

```{r volcano-plot, warning=FALSE, fig.width=10.5, fig.height=7.5}
for (nm in names(tests)) {
    plots <- plotVolcano(
        sce = sce, res = tests[[nm]], testType = stattest, 
        xv = xv, yv = yv, xvma = xvma, volcind = volcind, 
        plotnote = testres$plotnotes[[nm]], 
        plottitle = testres$plottitles[[nm]], 
        plotsubtitle = testres$plotsubtitles[[nm]],
        volcanoFeaturesToLabel = volcanoFeaturesToLabel, 
        volcanoMaxFeatures = volcanoMaxFeatures,
        baseFileName = sub("\\.Rmd$", "", knitr::current_input()), 
        comparisonString = nm, stringDb = string_db,
        featureCollections = featureCollections, 
        complexFDRThr = complexFDRThr,
        maxNbrComplexesToPlot = maxNbrComplexesToPlot,
        curveparam = testres$curveparams[[nm]],
        abundanceColPat = aval)
    print(plots$gg)
    if (!is.null(plots$ggma)) {
        print(plots$ggma)
    }
    interactiveVolcanos[[nm]] <- plots$ggint
}
```

```{r interactive-volcanos, echo = FALSE, eval = addInteractiveVolcanos, fig.width = 10.5, fig.height = 8}
interactiveVolcanos
```

For each comparison, we also save a text file with the "significant" features 
(defined as those colored in the volcano plots above). The features are 
ordered by the logFC value. In addition, we save a file with iBAQ values for 
all features that are "significant" in at least one comparison, across all the 
samples used in at least one comparison. 

```{r export-testres}
abundanceExport <- lapply(structure(names(tests), names = names(tests)), function(nm) {
    tests[[nm]] %>% dplyr::select(pid, showInVolcano, starts_with(aval))
})
for (nm in names(abundanceExport)) {
    colnames(abundanceExport[[nm]])[colnames(abundanceExport[[nm]]) == "showInVolcano"] <- 
        paste0("showInVolcano_", nm)
}
abundanceExport <- Reduce(function(...) dplyr::full_join(..., by = "pid", 
                                                    suffix = c("", "__remove")),
                     abundanceExport) %>%
    dplyr::select(pid, starts_with("showInVolcano"), matches("avg$"), 
                  matches("sd$"), everything()) %>%
    dplyr::select(-contains("__remove")) %>%
    dplyr::filter(if_any(starts_with("showInVolcano"), ~ .x))
write.table(abundanceExport, file = sub("\\.Rmd$", "_abundance_values_significant.txt",
                                   knitr::current_input()),
            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```


```{r merge-tests}
## Merge results from all tests and add to rowData(sce)
for (nm in names(tests)) {
    idx <- which(colnames(tests[[nm]]) != "pid")
    colnames(tests[[nm]])[idx] <- paste0(nm, ".", colnames(tests[[nm]])[idx])
}
all_tests <- as.data.frame(Reduce(function(...) dplyr::full_join(..., by = "pid"),
                                  tests), optional = TRUE)
rownames(all_tests) <- all_tests$pid
all_tests$pid <- NULL

stopifnot(rownames(sce) == rownames(all_tests))
SummarizedExperiment::rowData(sce) <-
  cbind(SummarizedExperiment::rowData(sce), all_tests)
```

The UpSet plot below shows the overlap among the "significant" proteins 
(defined as the ones that are colored in the volcano plots above, based 
on the user specifications) from the different comparisons. Note that if 
there are many comparisons, not all combinations may be displayed in 
the plot (only the 50 combinations with the largest number of proteins 
are shown, for interpretability reasons). Moreover, the UpSet plot will 
only be shown if at least two comparisons have been made, and there are 
at least two comparisons where any proteins were deemed significant.

```{r upset-tests, fig.width = 9}
tmpsign <- all_tests %>% dplyr::select(contains("showInVolcano")) + 0
tmpsign[is.na(tmpsign)] <- 0
colnames(tmpsign) <- gsub("\\.showInVolcano$", "", colnames(tmpsign))
if (length(tests) > 1 && sum(colSums(tmpsign) > 0) > 1) {
    UpSetR::upset(tmpsign, nsets = length(tests), nintersects = 50, 
                  order.by = "freq")
}
```

Finally, we display the top significant feature sets in each of the tested 
collections, for each comparison (if applicable). We recommend that the 
(adjusted) p-values for the feature sets is interpreted with caution, 
especially in situations where the protein abundances are measured on an 
isoform level and the feature sets are defined on the protein level, since 
there will be many (sometimes strongly correlated) features corresponding to 
a single gene or protein annotated to a feature set. 

```{r top-feature-sets, fig.height = 4 * length(featureCollections) + 0.1}
for (nm in names(testres$topsets)) {
    if (length(testres$topsets[[nm]]) > 0) {
        plts <- lapply(names(testres$topsets[[nm]]), function(snm) {
            df <- testres$topsets[[nm]][[snm]]
            if (nrow(df) > 10) df <- df[seq_len(10), ]
            if (nrow(df) > 0) {
                ggplot(df %>% dplyr::mutate(set = factor(.data$set, levels = rev(.data$set))), 
                       aes(x = .data$set, y = -log10(.data[[paste0(nm, "_FDR")]]))) + 
                    geom_bar(stat = "identity") + 
                    coord_flip() + theme_bw() + 
                    labs(x = "", title = snm) + 
                    theme(axis.text = element_text(size = 6))
            } else {
                NULL
            }
        })
        if (sum(sapply(plts, is.null)) < length(plts)) {
            print(cowplot::plot_grid(plotlist = plts, ncol = 1, align = "v"))
        }
    }
}
```


# Table with direct database links to sequences, functional information and predicted structures

The table below provides autogenerated links to the UniProt and 
AlphaFold pages (as well as selected organism-specific databases) for the 
majority protein IDs corresponding to each feature in the data set. 
UniProt is a resource of protein sequence and functional information 
hosted by EMBL-EBI, PIR and SIB. The AlphaFold Protein Structure Database, 
developed by DeepMind and EMBL-EBI, provides open access to protein structure 
predictions for the human proteome and other key proteins of interest. 
Note that (depending on the species) many proteins are not yet covered in 
AlphaFold (in this case, the link below will lead to a non-existent page). 

```{r linktable, warning = FALSE}
linkTable <- makeDbLinkTable(
    df = as.data.frame(rowData(sce)) %>%
        tibble::rownames_to_column("pid") %>%
        dplyr::select(pid, Majority.protein.IDs), 
    idCol = "Majority.protein.IDs", 
    speciesCommon = speciesInfo$speciesCommon, 
    addSpeciesSpecificColumns = TRUE, 
    convTablePomBase = pbconv, 
    convTableWormBase = wbconv
) %>%
    dplyr::select(-Majority.protein.IDs)
DT::datatable(as.data.frame(linkTable), escape = FALSE,
              options = list(scrollX = TRUE, pageLength = 20))
```


# Assemble SingleCellExperiment object

Finally, we assemble all the information calculated above in a 
[SingleCellExperiment](https://bioconductor.org/packages/SingleCellExperiment/) 
object, which can later be used e.g. for exploration with 
[iSEE](https://bioconductor.org/packages/iSEE/).

```{r make-sce}
sce <- prepareFinalSCE(
    sce = sce, baseFileName = sub("\\.Rmd$", "", knitr::current_input()),
    featureCollections = featureCollections, expType = "MaxQuant")

## Add experiment metadata
S4Vectors::metadata(sce) <- c(
    S4Vectors::metadata(sce), 
    list(
        mqFile = mqFile,
        aName = aName,
        iColPattern = iColPattern,
        imputeMethod = imputeMethod,
        ctrlGroup = ctrlGroup,
        allPairwiseComparisons = allPairwiseComparisons,
        normMethod = normMethod,
        stattest = stattest,
        minlFC = minlFC,
        analysisDate = as.character(Sys.Date()),
        rmdFile = knitr::current_input(dir = TRUE)
    )
)

sce
```


# Run PCA

We run a principal component analysis to obtain a reduced dimensionality 
representation of the data, in order to visualize the samples in two dimensions. 

```{r run-pca, dpi=100}
interactivePCAs <- htmltools::tagList()
for (a in unique(c(assayForTests, assayForPlots))) {
    sce <- scater::runPCA(sce, exprs_values = a, 
                          ncomponents = 10, ntop = Inf,
                          BSPARAM = BiocSingular::ExactParam(), 
                          name = paste0("PCA_", a))
    pcadf <- data.frame(sampleLabel = colnames(sce), 
                        reducedDim(sce, paste0("PCA_", a))[, 1:2], 
                        group = sce$group)
    interactivePCAs[[a]] <- plotly::ggplotly(
        ggplot(pcadf, aes(x = PC1, y = PC2, color = group, label = sampleLabel)) + 
            geom_point(alpha = 0.5, size = 5) + 
            ggtitle(paste0("Assay: ", a)) + 
            theme_cowplot(),
        width = 750, height = 500
    )
}
```

```{r interactive-pcas, echo = FALSE}
interactivePCAs
```

# Heatmap with hierarchical clustering

For another birds-eye view of the data, we represent it using a heatmap of 
the (imputed and normalized) log intensities, and cluster the samples and 
proteins using hierarchical clustering. In the first heatmap below, the values 
represent the normalized log intensities directly. In the second heatmap, 
the values for each protein have been centered to mean 0. The latter is also 
exported to a pdf file with row labels for further exploration. 

```{r heatmap, fig.height=8, message=FALSE}
gg_color_hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
}
groupcols = gg_color_hue(length(unique(sce$group)))
names(groupcols) <- levels(factor(sce$group))

ht <- ComplexHeatmap::Heatmap(
    assay(sce, assayForTests),
    name = assayForTests,
    show_row_names = FALSE,
    use_raster = TRUE,
    top_annotation = columnAnnotation(group = sce$group,
                                      col = list(group = groupcols))
)
draw(ht, merge_legends = TRUE)

## Centered
ht <- ComplexHeatmap::Heatmap(
    t(scale(t(assay(sce, assayForTests)), center = TRUE, scale = FALSE)),
    name = paste0(assayForTests, "\ncentered"),
    show_row_names = FALSE,
    use_raster = TRUE,
    top_annotation = columnAnnotation(group = sce$group,
                                      col = list(group = groupcols))
)
draw(ht, merge_legends = TRUE)
```

```{r save-heatmap, results="hide"}
## Save to pdf (show row names, but no row dendrogram, order samples by group)
pdf(sub("\\.Rmd$", "_heatmap_centered.pdf", knitr::current_input(dir = TRUE)),
    width = ncol(sce)/3.3 + 3.375, height = nrow(sce)/6.33 + 4)
plotassay <- assay(sce, assayForTests)
colnames(plotassay) <- sub(iColPattern, "", colnames(plotassay))
ht <- ComplexHeatmap::Heatmap(
    t(scale(t(plotassay), center = TRUE, scale = FALSE)),
    name = paste0(assayForTests, "\ncentered"),
    show_row_names = TRUE, show_row_dend = FALSE,
    cluster_columns = TRUE, use_raster = TRUE, column_split = sce$group,
    top_annotation = columnAnnotation(group = sce$group,
                                      col = list(group = groupcols)),
    bottom_annotation = columnAnnotation(group = sce$group,
                                         col = list(group = groupcols)),
    heatmap_legend_param = list(direction = "horizontal"),
    column_title_gp = gpar(fontsize = 6)
)
draw(ht, merge_legends = TRUE, heatmap_legend_side = "bottom", 
     annotation_legend_side = "bottom")
dev.off()
```


# Save SingleCellExperiment object

The `SingleCellExperiment` object created above is saved in the following location: 

```{r save-sce}
sceFile <- sub("\\.Rmd$", paste0("_sce.rds"), knitr::current_input(dir = TRUE))
saveRDS(sce, file = sceFile)
sceFile
```

In addition, all feature information (the `rowData` of the `SingleCellExperiment`)
is written to a text file:

```{r save-rowdata}
textFile <- sub("\\.Rmd$", paste0("_feature_info.txt"), 
                knitr::current_input(dir = TRUE))
write.table(as.data.frame(rowData(sce)) %>% 
                tibble::rownames_to_column("FeatureID"), 
            file = textFile, row.names = FALSE, col.names = TRUE,
            quote = FALSE, sep = "\t")
textFile
```


# Explore the data interactively

For interactive exploration of your results, we generate a script to launch 
an adapted [iSEE](https://bioconductor.org/packages/iSEE/) interface. The 
script can be sourced from an R console: 

```{r isee-script-path, echo=FALSE}
iSEEScript <- sub("\\.Rmd$", paste0("_iSEE.R"), knitr::current_input(dir = TRUE))
```


```{r create-markdown-chunks-dynamically, include=FALSE}
out <- paste0("\n```{r, results='asis', echo=TRUE, eval=FALSE, class.source = 'fold-show'}\n\nsource('", 
              iSEEScript, "')\n\n```")
```

`r paste(knitr::knit(text = out), collapse = '\n')`


That will open up an iSEE session where you can interactively explore your data.

```{r make-isee-script, echo=FALSE}
makeiSEEScript(iSEEScript = iSEEScript, sceFile = sceFile, 
               aName = aName, tests = tests, 
               assayForPlots = assayForTests,
               assayForHeatmaps = "iBAQ",
               includeFeatureSetTable = (length(featureCollections) > 0))
```

# Session info {#session-info}

This report was compiled with the following package versions:

<details>
<summary><b>
Click to expand
</b></summary>
```{r session-info, echo=FALSE}
sessionInfo()
```
</details>


# References


